---
title: "Antioxidant.Rmd"
author: "EL Strand"
date: "5/27/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Code modified from R. Cunning, H. Putnam, and E5 group; link [here](https://github.com/urol-e5/timeseries/blob/master/timepoint_1/scripts/antioxidant_capacity.Rmd).

9-10 antioxidant plates with each date containing a standard curve (not each plate will have a standard curve).  

Clears workspace. 
```{r}
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(broom)
library(Rmisc)
library(lme4)
library(car)
library(sjPlot)
```

Read in all data files. 
```{r}
# List protein data files
tac_path = "Physiology_variables/Total_Antioxidant_Capacity/"      # Path to prot data directory
all_tac_files <- list.files(path = tac_path, pattern = "*.csv")          # List all files in directory
tac_platemaps <- list.files(path = tac_path, pattern = "platemap")       # List platemap files
tac_data_files <- setdiff(all_tac_files, tac_platemaps)                 # List data files

# Read in all files into tibble
df <- tibble(file = tac_data_files) %>%
  separate(file, into = c("date", "plate", "readtime"), remove = FALSE) %>%
  unite(plate, date, plate) %>%                                                          # keep readtime separate so R can pair the map to both final and initial files 
  mutate(platemap = map(plate, ~read_csv(paste0(tac_path, ., "_TAC_platemap.csv"))),
         tac_data = map( file, ~read_csv(paste0(tac_path, .)))) 

# Merge platemap and data for each plate
df <- df %>%
  mutate(merged = map2(platemap, tac_data, ~ right_join(.x, .y)))


# files that have 'standard' or 'NA' are characters, but all numeric plug ids are numeric and need to be changed to character
# 3-6, 9-12, 17-18, 21-24

#df[[6]][[1:32]]$Plug_ID <- sapply(df[[6]][[1:32]]$Plug_ID,as.character)

# come back to a cleaner way to do this 
df[[6]][[3]]$Plug_ID <- as.character(df[[6]][[3]]$Plug_ID)
df[[6]][[4]]$Plug_ID <- as.character(df[[6]][[4]]$Plug_ID)
df[[6]][[5]]$Plug_ID <- as.character(df[[6]][[5]]$Plug_ID)
df[[6]][[6]]$Plug_ID <- as.character(df[[6]][[6]]$Plug_ID)
df[[6]][[9]]$Plug_ID <- as.character(df[[6]][[9]]$Plug_ID)
df[[6]][[10]]$Plug_ID <- as.character(df[[6]][[10]]$Plug_ID)
df[[6]][[11]]$Plug_ID <- as.character(df[[6]][[11]]$Plug_ID)
df[[6]][[12]]$Plug_ID <- as.character(df[[6]][[12]]$Plug_ID)
df[[6]][[17]]$Plug_ID <- as.character(df[[6]][[17]]$Plug_ID)
df[[6]][[18]]$Plug_ID <- as.character(df[[6]][[18]]$Plug_ID)
df[[6]][[21]]$Plug_ID <- as.character(df[[6]][[21]]$Plug_ID)
df[[6]][[22]]$Plug_ID <- as.character(df[[6]][[22]]$Plug_ID)
df[[6]][[23]]$Plug_ID <- as.character(df[[6]][[23]]$Plug_ID)
df[[6]][[24]]$Plug_ID <- as.character(df[[6]][[24]]$Plug_ID)
```

Plot standard curves and calculate fitted model values. 
```{r}
# Create standards following kit instructions
standards <- tribble(
  ~std, ~conc_mM,
  1,           1,
  2,         0.5,
  3,        0.25,
  4,       0.125,
  5,      0.0625,
  6,     0.03125,
  7,     0.01560,
  8,     0.00780,
  9,     0.00390,
 10,           0
)

std_curve <- df %>%
  unnest(merged) %>%
  filter(grepl("Standard", Plug_ID)) %>%
  dplyr::rename(absorbance = '490:490') %>% 
  select(plate, Well, Plug_ID, absorbance, readtime) %>%
  dplyr::rename(std = Plug_ID) %>%
  mutate(std = as.numeric(str_sub(std, 9, 10))) %>%
  spread(readtime, absorbance) %>%
  mutate(net_abs = final - initial) %>%
  mutate(net_abs = net_abs - net_abs[std == "10"]) %>%
  left_join(standards) %>%
  group_by(plate)

## Visualizing standard curves and choosing nov10 to use for all samples 
std_curve %>%
  ggplot(aes(x = net_abs, y = conc_mM, group = plate, color = plate)) +
  geom_point() + theme_bw() +
  labs(title = "Standard curve") 

nov10 <- std_curve %>% subset(plate=="20201110_plate1") 

nov10plot <- nov10 %>%
  ggplot(aes(x = net_abs, y = conc_mM, group = plate, color = plate)) +
  geom_point() + theme_bw() +
  #geom_line(aes(x = net_abs, y = .fitted)) +
  labs(title = "Nov 10 standard curve"); nov10plot

# Fit linear model
lmod <- lm(conc_mM ~ net_abs, data = nov10)

# Get fitted values
fitted <- lmod %>% 
  broom::augment(newdata = tibble(net_abs = seq(0, max(std_curve$net_abs), 0.005)))
```

Calculating TAC values. 
```{r}
tac <- df %>%
  unnest(merged) %>%
  filter(!grepl("Standard", Plug_ID)) %>%             # filters out the standard values 
  dplyr::rename(absorbance = '490:490') %>% 
  select(plate, Well, Plug_ID, absorbance, readtime) %>%
  spread(readtime, absorbance) %>%
  # Use standard curve to predict concentrations based on sample absorbance values
  mutate(net_abs = final - initial,
         uae.mmol.L = map_dbl(net_abs, ~ predict(lmod, newdata = data.frame(net_abs = .)))) 

nov10plot +
  geom_point(data = tac, aes(x = net_abs, y = uae.mmol.L), pch = "X", cex = 3, alpha = 0.3) +
    labs(title = "All samples projected on standard curve")

tac <- tac %>%
  group_by(Plug_ID) %>%
  mutate(uae.mmol.L = mean(uae.mmol.L)) %>% select(-Well, -final, -initial, -net_abs) %>% distinct()
```

Read in metadata and calculate TAC.
```{r}
# Protein output data = already has metadata and homogenate volume
protein <- read.csv("Output/soluble_protein.csv") %>% na.omit() %>% select(Plug_ID, vol_mL, Species, Tank, Timepoint, Temperature, Treatment, CO2, prot_ug)
protein$Plug_ID <- as.character(protein$Plug_ID)

# surface area 
sa <- read.csv("Physiology_variables/Respirometry/coral_surface_area.csv") %>% select(Sample_ID, surface.area.cm2) %>%
  dplyr::rename(Plug_ID = Sample_ID)
sa$Plug_ID <- as.character(sa$Plug_ID)

protein <- full_join(protein, sa, by = "Plug_ID")

# Join TAC data with metadata
tac <- left_join(tac, protein) %>%
  mutate(cre.umol.L = uae.mmol.L * 2189) %>%   # Convert to CRE (see product manual) per unit sample volume
  mutate(cre.umol = cre.umol.L * (vol_mL / 1000),  # Convert to CRE per coral by multiplying by homog. vol.
         cre.umol.mgprot = cre.umol / (prot_ug / 1000),
         cre.umol.cm2 = cre.umol/surface.area.cm2)  # Convert to CRE per mg protein by dividing by protein
### account for 25 uL prot and 20 uL TAC difference 

tac <- tac %>% filter(!is.na(Plug_ID)) %>% filter(!is.na(Species))

## Remove 1430 and 2391 as an outlier 
tac <- tac %>% filter(!grepl("1430", Plug_ID)) %>% filter(!grepl("2391", Plug_ID))

## data vis and outlier check 
boxplot(tac$cre.umol.mgprot)
hist(tac$cre.umol.mgprot)
boxplot.stats(tac$cre.umol.mgprot)$out

#looking for any plate that is an outlier 
tac %>% ggplot(aes(x=plate, y=cre.umol.mgprot)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  facet_grid(Species ~ Treatment) + theme_bw()

# List missing values. 
which(is.na(tac))

# taking out plate 
tac <- tac %>% select(-plate) %>% distinct()

# output data
tac %>%
  write_csv(., file = "Output/antioxidant.csv")
```

Calculating means. 
```{r}
tac.means <- tac %>% summarySE(measurevar="cre.umol.mgprot", groupvars=c("Timepoint", "Treatment", "Temperature", "CO2", "Species")) %>%
  filter(N>2) %>%
  mutate(Species = case_when(
    Species == "Mcapitata" ~ "Montipora capitata",
    Species == "Pacuta" ~ "Pocillopora acuta"
  )) %>% subset(!Timepoint == "Day 1" & !Timepoint == "Day 2")

tac.means$Timepoint <- factor(tac.means$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week","12 week","16 week"))

# Looking at sample sizes 
dplyr::filter(tac.means, N > 6) # this output should be zero b/c the biggest sample size is 6
dplyr::filter(tac.means, N < 6) # this output should just be the Day 1, Day 2, Week 1, 12 week, and 16 week decreased sample sizes

filter(tac, Timepoint == "6 week" & Species == "Mcapitata" & Treatment == "ATAC")

tac.cm2.means <- tac %>% summarySE(measurevar="cre.umol.cm2", groupvars=c("Timepoint", "Treatment", "Temperature", "CO2", "Species")) %>%
  filter(N>2) %>%
  mutate(Species = case_when(
    Species == "Mcapitata" ~ "Montipora capitata",
    Species == "Pacuta" ~ "Pocillopora acuta"
  )) %>% subset(!Timepoint == "Day 1" & !Timepoint == "Day 2")
tac.cm2.means$Timepoint <- factor(tac.cm2.means$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week","12 week","16 week"))
```
Figure construction.
```{r}
cols <- c("blue", "lightblue", "salmon", "red3")

tac.means <- tac.means %>% subset(!Timepoint == "Day 1" & !Timepoint == "Day 2")
tac.means.stress <- tac.means %>% subset(!Timepoint == "Day 1" & !Timepoint == "Day 2" & !Timepoint == "12 week" & !Timepoint == "16 week")

tac.plot <- ggplot(tac.means, aes(x=Timepoint, y = cre.umol.mgprot, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(~Species) +
  geom_errorbar(aes(ymin=cre.umol.mgprot-se, ymax=cre.umol.mgprot+se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Timepoint") + 
  theme_classic() + 
  ylab(expression("Copper Reducing Equivalents" ~(umol ~mg^-1))) +
  #ggtitle("Total Antioxidant Capacity") +
  theme(legend.position = "left") +
  geom_vline(xintercept = c(5.2), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 12, color = "black", face = "bold.italic"),
  # Change axis line
  axis.line = element_line(colour = "black")); tac.plot

ggsave(file="Output/Final_Figures/Antioxidant-full.pdf", tac.plot, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Antioxidant-full.png", tac.plot, width = 11, height = 6, units = c("in"))


tac.plot.stress <- ggplot(tac.means.stress, aes(x=Timepoint, y = cre.umol.mgprot, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(~Species) +
  geom_errorbar(aes(ymin=cre.umol.mgprot-se, ymax=cre.umol.mgprot+se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Timepoint") + 
  theme_classic() + 
  ylab(expression("CRE" ~(umol ~mg^-1))) +
  #ggtitle("Total Antioxidant Capacity") +
  theme(legend.position = "none") +
  #geom_vline(xintercept = c(5.2), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.text = element_text(size = 13)) +
  theme(legend.title = element_text(size = 15)) +
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 12, color = "black", face = "bold.italic"),
  # Change axis line
  axis.line = element_line(colour = "black")); tac.plot

ggsave(file="Output/Final_Figures/Antioxidant.png", tac.plot.stress, width = 7, height = 4, units = c("in"))

```

Species comparison
```{r}
tac.species <- tac %>% subset(!Timepoint == "Day 1" & !Timepoint == "Day 2" & !Timepoint == "12 week" & !Timepoint == "16 week")
tac.species.means <- tac.species %>% summarySE(measurevar="cre.umol.mgprot", groupvars=c("Species"))

Spp.tac.plot <- ggplot(tac.species.means, aes(x=Species, y = cre.umol.mgprot, group = Species)) + geom_point(size=3) + geom_line() +
  geom_errorbar(aes(ymin=cre.umol.mgprot-se, ymax=cre.umol.mgprot+se), width=.2) +
  theme_classic() + 
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  ylab(expression("CRE"~(umol ~mg^-1))) +
  theme(legend.position = "none") +
  theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))); Spp.tac.plot

ggsave(file="Output/Final_Figures/Antioxidant_spp.png", Spp.tac.plot, width = 5, height = 3, units = c("in"))
```

Publication figure 

```{r}
antioxidant_plot_full <- ggplot(tac.means, aes(x=Timepoint, y = cre.umol.mgprot, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(~Species, scales = "free") +
  geom_errorbar(aes(ymin=cre.umol.mgprot -se, ymax=cre.umol.mgprot +se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Timepoint") + 
  theme_classic() + 
  ylab(expression("CRE"~(umol ~mg^-1))) +
  theme(legend.position = c(0.6,0.7)) +
  geom_vline(xintercept = c(5.2), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 20, color = "black", face = "bold.italic"), # Change axis line
  axis.line = element_line(colour = "black")) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 13)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.title = element_text(size = 15)); antioxidant_plot_full # Change axis line

ggsave(file="Output/Supplemental_figures/fullantioxidant.png", antioxidant_plot_full, width = 11, height = 6, units = c("in"))
```

```{r}
ggplot(tac.cm2.means, aes(x=Timepoint, y = cre.umol.cm2, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(~Species, scales = "free") +
  geom_errorbar(aes(ymin=cre.umol.cm2 -se, ymax=cre.umol.cm2 +se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Timepoint") + 
  theme_classic() + 
  ylab(expression("CRE"~(umol ~cm^-2))) +
  theme(legend.position = c(0.6,0.7)) +
  geom_vline(xintercept = c(5.2), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 20, color = "black", face = "bold.italic"), # Change axis line
  axis.line = element_line(colour = "black")) + theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 13)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  theme(legend.title = element_text(size = 15))
```



# Statistics 

### Host Total Antioxidant Capacity

Read in TAC data file so that you don't have to re-run the entire script to do the statistics. 
```{r}
antioxidant <- read.csv("Output/antioxidant.csv")

mcap.antiox <- antioxidant %>% subset(Species == "Mcapitata") %>% subset(Timepoint == "1 week" | Timepoint == "2 week" | Timepoint == "4 week" | Timepoint == "6 week" | Timepoint == "8 week")

mcap.recovery.antiox <- antioxidant %>% subset(Species == "Mcapitata") %>% subset(Timepoint == "12 week" | Timepoint == "16 week")

pacuta.antiox <- antioxidant %>% subset(Species == "Pacuta") %>% subset(Timepoint == "1 week" | Timepoint == "2 week" | Timepoint == "4 week" | Timepoint == "6 week" | Timepoint == "8 week")

hist(mcap.antiox$cre.umol.mgprot) 
hist(pacuta.antiox$cre.umol.mgprot) 
hist(spp.antiox$cre.umol.mgprot) 
hist(mcap.recovery.antiox$cre.umol.mgprot) 
```

Mixed effect modeling  

```{r}
TAC_LMER <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2*Species + (1|Tank), na.action=na.omit, data=spp.antiox) 

Pacuta_TAC_LMER <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank:Temperature:Timepoint) + (1|Tank:CO2:Timepoint), na.action=na.omit, data=pacuta.antiox) 
Pacuta_TAC_LMER2 <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank:Temperature:Timepoint) + (1|Tank:CO2:Timepoint), na.action=na.omit, data=pacuta.antiox, contrasts = list(Timepoint = "contr.sum", Temperature = "contr.sum", CO2 = "contr.sum")) 

Mcap_TAC_LMER <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank:Temperature:Timepoint) + (1|Tank:CO2:Timepoint), na.action=na.omit, data=mcap.antiox) 
Mcap_TAC_LMER2 <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank:Temperature:Timepoint) + (1|Tank:CO2:Timepoint), na.action=na.omit, data=mcap.antiox, contrasts = list(Timepoint = "contr.sum", Temperature = "contr.sum", CO2 = "contr.sum")) 

Mcap_TAC_rec_LMER <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank:Temperature:Timepoint) + (1|Tank:CO2:Timepoint), na.action=na.omit, data=mcap.recovery.antiox) 
Mcap_TAC_rec_LMER2 <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank:Temperature:Timepoint) + (1|Tank:CO2:Timepoint), na.action=na.omit, data=mcap.recovery.antiox, contrasts = list(Timepoint = "contr.sum", Temperature = "contr.sum", CO2 = "contr.sum")) 

qqPlot(residuals(Pacuta_TAC_LMER)) 
qqPlot(residuals(Pacuta_TAC_LMER2)) 
qqPlot(residuals(Mcap_TAC_LMER)) 
qqPlot(residuals(Mcap_TAC_LMER2))
qqPlot(residuals(Mcap_TAC_rec_LMER)) 
qqPlot(residuals(Mcap_TAC_rec_LMER2))

hist(residuals(Pacuta_TAC_LMER)) 
hist(residuals(Pacuta_TAC_LMER2)) 
hist(residuals(Mcap_TAC_LMER)) 
hist(residuals(Mcap_TAC_LMER2))
hist(residuals(Mcap_TAC_rec_LMER)) 
hist(residuals(Mcap_TAC_rec_LMER2))

## Type III Wald chisquare tests
pacuta.tac.anova <- Anova(Pacuta_TAC_LMER, ddf="lme4", type='III')
pacuta.tac.anova2 <- Anova(Pacuta_TAC_LMER2, ddf="lme4", type='III')

mcap.tac.anova <- Anova(Mcap_TAC_LMER, ddf="lme4", type='III')
mcap.tac.anova2 <- Anova(Mcap_TAC_LMER2, ddf="lme4", type='III')

mcap.tac.rec.anova <- Anova(Mcap_TAC_rec_LMER, ddf="lme4", type='III')
mcap.tac.rec.anova2 <- Anova(Mcap_TAC_rec_LMER2, ddf="lme4", type='III')

pacuta.tac.anova
pacuta.tac.anova2
mcap.tac.anova
mcap.tac.anova2
mcap.tac.rec.anova
mcap.tac.rec.anova2

capture.output(pacuta.tac.anova2, file = "Output/Statistics/Stress_statistics/TAC/TAC_Pacuta_sums.txt")
capture.output(mcap.tac.anova2, file = "Output/Statistics/Stress_statistics/TAC/TAC_Mcap_sums.txt")
capture.output(mcap.tac.rec.anova2, file = "Output/Statistics/Recovery_statistics/TAC_Mcap_recovery_sums.txt")
```

#### Species comparison 

```{r}
spp.antiox <- antioxidant %>% subset(Timepoint == "1 week")

spp.tac.t <- t.test(cre.umol.mgprot~Species, data = spp.antiox)

capture.output(spp.tac.t, file = "Output/Statistics/Species_comparison/TAC.txt")

Spp.tac.plot <- ggplot(spp.antiox, aes(x=Species, y = cre.umol.mgprot, group = Species)) + geom_point(size=3) + geom_boxplot() +
  theme_classic() + 
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  ylab(expression("CRE"~(umol ~mg^-1))) +
  theme(legend.position = "none") +
  theme(axis.text = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))); Spp.tac.plot

ggsave(file="Output/Final_Figures/Antioxidant_spp.png", Spp.tac.plot, width = 5, height = 3, units = c("in"))
```


