---
title: "CarbChem"
author: "EL Strand"
date: "9/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Contents:
1. [Experimental Tank Conditions](#exp_tank) 


Clears workspace. 
```{r}
rm(list=ls())
```

Load in required libraries.
```{r}
if("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if("seacarb" %in% rownames(installed.packages()) == 'FALSE') install.packages('seacarb') 
if("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 
if("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 

library(lubridate)
library(tidyverse)
library(seacarb) 
library(vegan)
library(emmeans)
library(gridExtra)
library(plotrix)
library(Rmisc)
```

##### <a name= ="exp_tank"></a> Experimental Tank Conditions ##### 
Discrete pH calculations from Tris calibrations.
```{r}
path <-("Environmental_data/Calibration_Files/Tris_calibrations") #set path to calibration files
file.names<-list.files(path = path, pattern = "csv$") #list all the file names in the folder to get only get the csv files
pH.cals <- data.frame(matrix(NA, nrow=length(file.names), ncol=3, dimnames=list(file.names,c("Date", "Intercept", "Slope")))) #generate a 3 column dataframe with specific column names

for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Calib.Data <-read.table(file.path(path,file.names[i]), header=TRUE, sep=",", na.string="NA", as.is=TRUE) #reads in the data files
  file.names[i]
  model <-lm(mVTris ~ TTris, data=Calib.Data) #runs a linear regression of mV as a function of temperature
  coe <- coef(model) #extracts the coeffecients
  summary(model)$r.squared #extracts the r squared
  plot(Calib.Data$mVTris, Calib.Data$TTris) #plots the regression data
  pH.cals[i,2:3] <- coe #inserts coefficients in the dataframe
  pH.cals[i,1] <- substr(file.names[i],1,8) #stores the file name in the Date column
}
colnames(pH.cals) <- c("Calib.Date",  "Intercept",  "Slope") #rename columns
pH.cals #view data

#constants for use in pH calculation 
R <- 8.31447215 #gas constant in J mol-1 K-1 
F <-96485.339924 #Faraday constant in coulombs mol-1
```

Read in probe measurements of pH, temperature from tanks.
```{r}
daily <- read.csv("Environmental_data/Daily_Temp_pH_Sal.csv", header=TRUE, sep=",", na.strings="NA") #load data with a header, separated by commas, with NA as NA

#merge with Seawater chemistry file
SW.chem <- merge(pH.cals, daily, by="Calib.Date")
```

Calculate total pH
```{r}
mvTris <- SW.chem$Temperature*SW.chem$Slope+SW.chem$Intercept #calculate the mV of the tris standard using the temperature mv relationships in the measured standard curves 
STris<-35 #salinity of the Tris
phTris<- (11911.08-18.2499*STris-0.039336*STris^2)*(1/(SW.chem$Temperature+273.15))-366.27059+ 0.53993607*STris+0.00016329*STris^2+(64.52243-0.084041*STris)*log(SW.chem$Temperature+273.15)-0.11149858*(SW.chem$Temperature+273.15) #calculate the pH of the tris (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
SW.chem$pH.Total<-phTris+(mvTris/1000-SW.chem$pH.MV/1000)/(R*(SW.chem$Temperature+273.15)*log(10)/F) #calculate the pH on the total scale (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
```

Quality check calculated tank conditions. This is helpful to run while the experiment is on-going to catch any typos in the datasheet or mis-typed equations.
```{r}
boxplot(SW.chem$Temperature)
range(na.omit(SW.chem$Temperature))

boxplot(SW.chem$pH.Total~SW.chem$Treatment)
min(na.omit(SW.chem$pH.Total))
which(SW.chem$pH.Total == min(na.omit(SW.chem$pH.Total)))

boxplot(SW.chem$Salinity)
range(na.omit(SW.chem$Salinity))
```

Removing acclimation and testing periods. 
```{r}
#remove CO2 and heat test tanks
SW.chem <- SW.chem %>%
  filter(Treatment != "Heat_Test") %>%
  filter(Treatment != "CO2_Test")

unique(as.character(SW.chem$Treatment))

#remove acclimation period
Trt.SW.chem  <- SW.chem %>%
  filter(Treatment != "Acclimation")
```

Plot Temp and pH for all daily measures.
```{r}
#arrange to longer format
Trt.SW.chem <- pivot_longer(Trt.SW.chem, c(Temperature, pH.Total)) %>% na.omit(Trt.SW.chem)

pdf("Output/Daily.Chem.pdf", width=10, height=4)
Trt.SW.chem %>%
  group_by(Date, Treatment) %>%
  ggplot(aes(x=Treatment, y=value)) +
  geom_boxplot() +
  facet_grid(name ~ ., scales="free_y")+
  theme_classic()

Trt.SW.chem.means <- summarySE(Trt.SW.chem, measurevar="value", groupvars=c("Treatment", "name"))
dev.off()
```
