---
title: "CarbChem"
author: "EL Strand"
date: "9/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clears workspace. 
```{r}
rm(list=ls())
```

Load in required libraries.
```{r}
if("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if("seacarb" %in% rownames(installed.packages()) == 'FALSE') install.packages('seacarb') 
if("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 
if("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 

library(lubridate)
library(tidyverse)
library(seacarb) 
library(vegan)
library(emmeans)
library(gridExtra)
library(plotrix)
library(Rmisc)
```

### DISCRETE TANK MEASUREMENTS ### 

Discrete pH calculations from Tris calibrations.
```{r}
path <-("Environmental_data/Calibration_Files/Tris_calibrations") #set path to calibration files
file.names<-list.files(path = path, pattern = "csv$") #list all the file names in the folder to get only get the csv files
pH.cals <- data.frame(matrix(NA, nrow=length(file.names), ncol=3, dimnames=list(file.names,c("Date", "Intercept", "Slope")))) #generate a 3 column dataframe with specific column names

for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Calib.Data <-read.table(file.path(path,file.names[i]), header=TRUE, sep=",", na.string="NA", as.is=TRUE) #reads in the data files
  file.names[i]
  model <-lm(mVTris ~ TTris, data=Calib.Data) #runs a linear regression of mV as a function of temperature
  coe <- coef(model) #extracts the coeffecients
  summary(model)$r.squared #extracts the r squared
  #plot(Calib.Data$mVTris, Calib.Data$TTris) #plots the regression data
  pH.cals[i,2:3] <- coe #inserts coefficients in the dataframe
  pH.cals[i,1] <- substr(file.names[i],1,8) #stores the file name in the Date column
}
colnames(pH.cals) <- c("Calib.Date",  "Intercept",  "Slope") #rename columns
pH.cals #view data

#constants for use in pH calculation 
R <- 8.31447215 #gas constant in J mol-1 K-1 
F <- 96485.339924 #Faraday constant in coulombs mol-1
```

Read in probe measurements of pH, temperature from tanks.
```{r}
daily <- read.csv("Environmental_data/Daily_Temp_pH_Sal.csv", header=TRUE, sep=",", na.strings="NA") #load data with a header, separated by commas, with NA as NA

#merge with Seawater chemistry file
SW.chem <- merge(pH.cals, daily, by="Calib.Date")
```

Calculate total pH. 
```{r}
mvTris <- SW.chem$Temperature*SW.chem$Slope+SW.chem$Intercept #calculate the mV of the tris standard using the temperature mv relationships in the measured standard curves 
STris<-35 #salinity of the Tris
phTris<- (11911.08-18.2499*STris-0.039336*STris^2)*(1/(SW.chem$Temperature+273.15))-366.27059+ 0.53993607*STris+0.00016329*STris^2+(64.52243-0.084041*STris)*log(SW.chem$Temperature+273.15)-0.11149858*(SW.chem$Temperature+273.15) #calculate the pH of the tris (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
SW.chem$pH.Total<-phTris+(mvTris/1000-SW.chem$pH.MV/1000)/(R*(SW.chem$Temperature+273.15)*log(10)/F) #calculate the pH on the total scale (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
```

Quality check calculated tank conditions. This is helpful to run while the experiment is on-going to catch any typos in the datasheet or mis-typed equations.
```{r}
par(mfrow=c(1,3))
boxplot(SW.chem$Temperature)
range(na.omit(SW.chem$Temperature))

boxplot(SW.chem$pH.Total~SW.chem$Treatment)
min(na.omit(SW.chem$pH.Total))
which(SW.chem$pH.Total == min(na.omit(SW.chem$pH.Total)))

boxplot(SW.chem$Salinity)
range(na.omit(SW.chem$Salinity))
```

Plotting Temperature
```{r}
SW.chem$Timepoint <- as.Date(SW.chem$Timepoint, format = "%m/%d/%y")

all.colors <- c("black", "lightblue", "blue", "salmon", "red3")

annotation <- data.frame(x = c(as.numeric(as.Date("2018-09-22")),as.numeric(as.Date("2018-09-29")),
         as.numeric(as.Date("2018-10-06")),as.numeric(as.Date("2018-10-20")),
         as.numeric(as.Date("2018-11-03")),as.numeric(as.Date("2018-11-17")),
         as.numeric(as.Date("2018-12-14")),as.numeric(as.Date("2019-01-10"))),
   y = c(22,22,22,22,22,22,22,22),
   label = c("Day1", "1 week", "2 week", "4 week", "6 week", "8 week", "12 week", "16 week"))

Discrete.Temp <- SW.chem %>%
  filter(Period2 != "Empty") %>%
  filter(Period2 != "Testing") %>%
  ggplot(aes(x = Timepoint, y = Temperature, color = Treatment)) +
  scale_color_manual(values = all.colors) +
  labs(x = "", y = "Temperature °C") +
  geom_jitter(width = 0.1) +
  geom_vline(xintercept = as.numeric(as.Date("2018-11-18")), colour="black", linetype="dotted") +
  theme_classic()+
  scale_x_continuous(breaks = c(17796, 17803, 17810, 17824, 17838, 17852, 17879, 17906), labels = c("Day1", "1 week", "2 week", "4 week", "6 week", "8 week", "12 week", "16 week"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave(file="Output/Discrete.Temp.pdf", Discrete.Temp, width = 11, height = 10, units = c("in"))
ggsave(file="Output/Discrete.Temp.png", Discrete.Temp, width = 11, height = 10, units = c("in"))

Discrete.pH <- SW.chem %>%
  filter(Period2 != "Empty") %>%
  filter(Period2 != "Testing") %>%
  ggplot(aes(x = Timepoint, y = pH.Total, color = Treatment)) +
  scale_color_manual(values = all.colors) +
  labs(x = "", y = "pH total scale") +
  geom_jitter(width = 0.1) +
  geom_vline(xintercept = as.numeric(as.Date("2018-11-18")), colour="black", linetype="dotted") +
  theme_classic()+
  scale_x_continuous(breaks = c(17796, 17803, 17810, 17824, 17838, 17852, 17879, 17906), labels = c("Day1", "1 week", "2 week", "4 week", "6 week", "8 week", "12 week", "16 week"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave(file="Output/Discrete.pH.pdf", Discrete.pH, width = 11, height = 10, units = c("in"))
ggsave(file="Output/Discrete.pH.png", Discrete.pH, width = 11, height = 10, units = c("in"))

Discrete.Temp
Discrete.pH
```

### DISCRETE TA CALCULATIONS ### 

Reading in TA cumulative file. 
```{r}
TA <- read.csv("Environmental_data/Total_Alkalinity/Cumulative_TA_Output.csv", header=TRUE, sep=",", na.strings="NA")  #read in  TA results
TA <- subset(TA , Type =="Tank")

##### SEAWATER CHEMISTRY ANALYSIS FOR DISCRETE MEASUREMENTS#####
#Seawater chemistry table from simultaneous TA, pH, temperature and salinity measurements
#merge calculated pH and daily measures with TA data and run seacarb
SW.chem$Sample.ID <- paste(SW.chem$Date, SW.chem$Sample.ID, sep='_') #generate new row with concatenated sample id
SW.chem <- subset(SW.chem , TA.Run =="Yes")
SW.chem <- merge(SW.chem,TA, by="Sample.ID", all = FALSE, sort = T) #merge seawater chemistry with total alkalinity

#Calculate CO2 parameters using seacarb
carb.output <- carb(flag=8, var1=SW.chem$pH.Total, var2=SW.chem$TA/1000000, S= SW.chem$Salinity, T=SW.chem$Temperature, P=0, Pt=0, Sit=0, pHscale="T", kf="pf", k1k2="l", ks="d") #calculate seawater chemistry parameters using seacarb
carb.output$ALK <- carb.output$ALK*1000000 #convert to µmol kg-1
carb.output$CO2 <- carb.output$CO2*1000000 #convert to µmol kg-1
carb.output$HCO3 <- carb.output$HCO3*1000000 #convert to µmol kg-1
carb.output$CO3 <- carb.output$CO3*1000000 #convert to µmol kg-1
carb.output$DIC <- carb.output$DIC*1000000 #convert to µmol kg-1
carb.output <- carb.output[,-c(1,4,5,8,10:13,19)] #subset variables of interest
carb.output <- cbind(SW.chem$Date,  SW.chem$Sample.ID,  SW.chem$Treatment,SW.chem$Timepoint, SW.chem$Period1, SW.chem$Period2, SW.chem$Type.x, carb.output) #combine the sample information with the seacarb output
colnames(carb.output) <- c("Date",  "Sample.ID",  "Treatment", "Timepoint","Period1", "Period2", "Type",	"Salinity", "Temperature", "pH",	"CO2",	"pCO2","HCO3",	"CO3",	"DIC", "TA",	"AragSat") #Rename columns to describe contents
carb.output$Tank <- substring(carb.output$Sample.ID, 10)
carb.output$Tank <- factor(carb.output$Tank)

write.table(carb.output, "Environmental_data/Seawater_chemistry_table_Output_All.csv", sep=",", row.names = FALSE) #save data
```

Subsetting to groups we want to visualize. 
```{r}
carb.output.long <- pivot_longer(carb.output, cols = Temperature:AragSat) %>%
  dplyr::rename(Measurement = name)

carb.output.long %>%
  filter(Treatment == "Acclimation" & Period2 == "Acclimation") %>%
  ggplot(aes(x = Timepoint, y = value, color = Treatment)) +
  scale_color_manual(values = all.colors) +
  facet_wrap(Measurement ~ ., scales="free_y") +
  labs(x = "", y = "Temperature °C") +
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black"))

trt.colors <- c("lightblue", "blue", "salmon", "red3")

#Acute Period
carb.output.long %>%
  filter(Period2 == "Acute") %>%
  ggplot(aes(x = Timepoint, y = value, color = Treatment)) +
  scale_color_manual(values = trt.colors) +
  facet_wrap(Measurement ~ ., scales="free_y") +
  labs(x = "", y = "Temperature °C") +
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black")) 

#Chronic Period
carb.output.long %>%
  filter(Period2 == "Chronic") %>%
  ggplot(aes(x = Timepoint, y = value, color = Treatment)) +
  scale_color_manual(values = trt.colors) +
  facet_wrap(Measurement ~ ., scales="free_y") +
  labs(x = "", y = "Temperature °C") +
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black")) 

#Recovery Period
carb.output.long %>%
  filter(Period2 == "Recovery") %>%
  ggplot(aes(x = Timepoint, y = value, color = Treatment)) +
  scale_color_manual(values = trt.colors) +
  facet_wrap(Measurement ~ ., scales="free_y") +
  labs(x = "", y = "Temperature °C") +
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black")) 

#All Treatment
Carb.Chem.Figs <- carb.output.long %>%
  ggplot(aes(x = Timepoint, y = value, color = Treatment)) +
  scale_color_manual(values = all.colors) +
  facet_wrap(Measurement ~ ., scales="free_y") +
  labs(x = "", y = "Temperature °C") +
  geom_jitter(width = 0.1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  axis.line = element_line(colour = "black")) +
  geom_vline(xintercept = as.numeric(as.Date("2018-11-18")), colour="black", linetype="dotted")+
  scale_x_continuous(breaks = c(17796, 17803, 17810, 17824, 17838, 17852, 17879, 17906), labels = c("Day1", "1 week", "2 week", "4 week", "6 week", "8 week", "12 week", "16 week"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave(file="Output/Carbonate-Chemistry.pdf", Carb.Chem.Figs, width = 11, height = 10, units = c("in"))
ggsave(file="Output/Carbonate-Chemistry.png", Carb.Chem.Figs, width = 11, height = 10, units = c("in"))

```
Generating Carb Chem Tables
```{r}

carb.output.long <- carb.output.long %>%
  filter(Period2 != "Empty") %>%
  filter(Period2 != "Testing")

carb.chem.DS <-  carb.output.long %>%
                group_by(Measurement, Period2, Treatment) %>%
                dplyr::summarize(mean=mean(value),
                          sem=(sd(value)/sqrt(n())),
                          .groups = "keep")

carb.chem.DS <- na.omit(carb.chem.DS)

carb.chem.DS <- data.frame(carb.chem.DS)
  
write.csv(carb.chem.DS, file="Output/Carbonate-Chemistry.csv")
          
```