---
title: "Cell Counts"
author: "EL Strand"
date: "5/24/2021"
output: html_document
---

Code modified from R. Cunning, H. Putnam, and E5 group; link [here](https://github.com/urol-e5/timeseries/blob/master/timepoint_1/scripts/Sym-Density.Rmd).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages.
```{r}
rm(list=ls())

library(plyr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggpubr)
library(purrr)
```

# Holobiont Integration sample processing 

## Loading meta data frames. 

```{r}
# Surface area data
sa <- read.csv("Physiology_variables/Respirometry/coral_surface_area.csv") %>% dplyr::rename(Plug_ID = Sample_ID) %>% select(Plug_ID, surface.area.cm2)
sa$Plug_ID <- as.character(sa$Plug_ID)

# Tissue homogenate volume data
homog_vols <- read_csv("Physiology_variables/homogenate_vol.csv") %>% select(Species, Plug_ID, vol_mL)
homog_vols$Plug_ID <- as.character(homog_vols$Plug_ID)

# Coral sample metadata
metadata <- read_csv("Environmental_data/Master_Fragment.csv") %>% select(Species, Plug_ID, Tank, Treatment, Temperature, CO2, Timepoint, ANALYSIS, Sample.Date) %>%
  subset(ANALYSIS=="Physiology" | ANALYSIS=="Physiology/Molecular")
metadata$Plug_ID <- as.character(metadata$Plug_ID)

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols) %>% full_join(sa) %>% filter(!is.na(vol_mL)) %>% 
  filter(!is.na(ANALYSIS))
```

## Import cellometer and haemocytometer data frames. 

```{r}
# Haemocytometer count data
haemo.raw <- read_csv("Physiology_variables/cell-counts/cell-counts-testing-haemo.csv") %>% filter(Tech == "Emma Strand")
  haemo.raw$Plug_ID <- as.character(haemo.raw$Plug_ID)
cellometer.raw <- read_csv("Physiology_variables/cell-counts/cell-counts-testing-cellometer.csv") %>% filter(Tech == "Emma Strand")
  cellometer.raw$Plug_ID <- as.character(cellometer.raw$Plug_ID)
```

## Calculate mean counts for both haemocytometer and cellometer.

### Haemocytometer

```{r}
# Calculate mean counts for each sample
haemo <- haemo.raw %>%
  select(Haemocytometer_Date, Homogenization, Plug_ID, Squares_Counted, matches("Count[0-9]")) %>%
  gather("rep", "count", -Haemocytometer_Date, -Homogenization, -Plug_ID, -Squares_Counted) %>% filter(!is.na(count)) %>%
  dplyr::group_by(Plug_ID, Homogenization) %>%
  mutate(count.mean = mean(count))

# Join mean counts with sample metadata
haemo <- full_join(haemo, metadata) 

# Normalize counts by homogenate volume and surface area
haemo <- haemo %>%
  mutate(cells.mL = count.mean * 10000 / Squares_Counted,
         cells = cells.mL * vol_mL,
         haemo.cells.cm2 = cells / surface.area.cm2)

haemo <- haemo %>% filter(!is.na(haemo.cells.cm2))
haemo$Timepoint <- factor(haemo$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week","12 week","16 week"))
```

### Cellometer

```{r}
cellometer <- cellometer.raw %>% 
  select(Homogenization, Plug_ID, matches("Cell_mL[0-2]")) %>%
  gather("rep", "count", -Homogenization, -Plug_ID) %>% filter(!is.na(count)) %>%
  dplyr::group_by(Plug_ID, Homogenization) %>%
  mutate(count.mean = mean(count))

cellometer <-full_join(cellometer, metadata) %>% filter(!is.na(Species)) %>% filter(!is.na(Homogenization))

cellometer <- cellometer %>% spread(rep, count) %>%
  mutate(cells.mL = count.mean,
         cells = cells.mL * vol_mL,
         cellometer.cells.cm2 = cells / surface.area.cm2) %>%
  mutate(cells1.mL = Cell_mL1,
         cells1 = cells1.mL * vol_mL,
         cellometer.cells1.cm2 = cells1 / surface.area.cm2) %>%
  mutate(cells2.mL = Cell_mL2,
         cells2 = cells2.mL * vol_mL,
         cellometer.cells2.cm2 = cells2 / surface.area.cm2)

cellometer$Timepoint <- factor(cellometer$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week","12 week","16 week"))
```

## Count1 vs Count2 comparison

### Create data frame and run model
```{r}
cellometer.c1c2.comparison <- cellometer %>% select(Plug_ID, Species, Cell_mL1, Cell_mL2, Homogenization)
  
c1c2.lm <- cellometer.c1c2.comparison %>%
  group_by(Homogenization, Species) %>%
  nest() %>% # nest to do this for each group (Species, Period) defined above 
  mutate(Mod = map(data, ~lm(Cell_mL1 ~ Cell_mL2, data = .x))) %>% # Fit linear model
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.squared, 3))) %>%   # Get the R2
  unnest(cols = c(data)) # unnest to plot below

c1c2 <- cellometer.c1c2.comparison %>%
  ggplot(aes(x=Cell_mL1, y=Cell_mL2, group = Homogenization, color = Homogenization)) + 
  geom_jitter(width = 0.1) + 
  facet_grid(Species ~ Homogenization) +
  labs(color='Period') + 
  theme_classic() +
  ylim(0, 1500000) + 
  xlim(0, 1500000) + 
  geom_abline(slope=1) + 
  geom_label(data = c1c2.lm, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1); c1c2

ggsave(file="Physiology_variables/cell-counts/Cellometer-c1c2.png", c1c2, width = 11, height = 6, units = c("in"))
```

## Cellometer and haemocytometer comparison

### Creating large dataframe.

```{r}
cellometer.comparison <- cellometer %>%
  select(Homogenization, Plug_ID, cellometer.cells.cm2, Species, cellometer.cells1.cm2, cellometer.cells2.cm2) 
haemo.comparison <- haemo %>% select(Homogenization, Plug_ID, haemo.cells.cm2)

## come back to 1048 -- this is full joining both of these 
method.comparison <- full_join(cellometer.comparison, haemo.comparison, join_by = "Plug_ID") %>% 
  filter(!is.na(Species)) %>%
  summarise_each(funs(first(na.omit(.)))) # combining rows that have the same data (plug ids with cellometer and haemcy but originally on two different rows)

mcap.comparison.samples <- method.comparison %>% subset(Species == "Mcapitata" & Homogenization == "1min") %>% na.omit()
nrow(mcap.comparison.samples)

pacuta.comparison.samples <- method.comparison %>% subset(Species == "Pacuta" & Homogenization == "1min") %>% na.omit()
nrow(pacuta.comparison.samples)
```

### Creating models for regression between cellometer and haemocytometer.

```{r}
methods.lm.mean <- method.comparison %>%
  group_by(Species, Homogenization) %>%
  nest() %>% # nest to do this for each group (Species, Homogenization) defined above 
  mutate(Mod = map(data, ~lm(cellometer.cells.cm2 ~ haemo.cells.cm2, data = .x))) %>% # Fit linear model
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.squared, 3))) %>%   # Get the R2
  unnest(cols = c(data)) # unnest to plot below 

methods.lm.count1 <- method.comparison %>%
  group_by(Species, Homogenization) %>%
  nest() %>% # nest to do this for each group (Species, Homogenization) defined above 
  mutate(Mod = map(data, ~lm(cellometer.cells1.cm2 ~ haemo.cells.cm2, data = .x))) %>% # Fit linear model
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.squared, 3))) %>%   # Get the R2
  unnest(cols = c(data)) # unnest to plot below 

methods.lm.count2 <- method.comparison %>%
  group_by(Species, Homogenization) %>%
  nest() %>% # nest to do this for each group (Species, Homogenization) defined above 
  mutate(Mod = map(data, ~lm(cellometer.cells2.cm2 ~ haemo.cells.cm2, data = .x))) %>% # Fit linear model
  mutate(R2 = map_dbl(Mod, ~round(summary(.x)$r.squared, 3))) %>%   # Get the R2
  unnest(cols = c(data)) # unnest to plot below 
```

### Plotting regressions for methods comparison. 

```{r}
methods.mean <- method.comparison %>%
  ggplot(aes(x=cellometer.cells.cm2, y=haemo.cells.cm2, group = Homogenization, color = Homogenization)) + 
  geom_jitter(width = 0.1) + 
  facet_grid(Species ~ Homogenization) +
  labs(color='Homogenization') +
  theme_classic() + 
  ylim(0, 3000000) + 
  xlim(0, 3000000) +
  ggtitle("Mean cellometer value") +
  geom_abline(slope=1) +
  geom_label(data = methods.lm.mean, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1); methods.mean

methods.count1 <- method.comparison %>%
  ggplot(aes(x=cellometer.cells1.cm2, y=haemo.cells.cm2, group = Homogenization, color = Homogenization)) + 
  geom_jitter(width = 0.1) + 
  facet_grid(Species ~ Homogenization) +
  labs(color='Homogenization') +
  theme_classic() + 
  ylim(0, 3000000) + 
  xlim(0, 3000000) +
  ggtitle("Count1 cellometer value") +
  geom_abline(slope=1) +
  geom_label(data = methods.lm.count1, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1); methods.count1

methods.count2 <- method.comparison %>%
  ggplot(aes(x=cellometer.cells2.cm2, y=haemo.cells.cm2, group = Homogenization, color = Homogenization)) + 
  geom_jitter(width = 0.1) + 
  facet_grid(Species ~ Homogenization) +
  labs(color='Homogenization') +
  theme_classic() + 
  ylim(0, 3000000) + 
  xlim(0, 3000000) +
  ggtitle("Count2 cellometer value") +
  geom_abline(slope=1) +
  geom_label(data = methods.lm.count2, 
             aes(x = Inf, y = Inf, 
                 label = paste("R2 = ", R2, sep = " ")),
             hjust = 1, vjust = 1); methods.count2
  
ggsave(file="Output/Cellometer-methodscompare-mean.png", methods.mean, width = 11, height = 6, units = c("in"))
```


```{r}
cols <- c("lightblue", "blue", "pink", "red")

haemo %>% na.omit() %>%
  ggplot(aes(x = Timepoint, y = haemo.cells.cm2, group = Treatment, color = Treatment)) +
  scale_color_manual(values = cols) +
  theme_classic() + 
  facet_grid(~Species) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  labs(x = "",y = "Cell Density (Cells/cm2)") +
  geom_jitter(width = 0.1) +  
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.1) +
  stat_summary(fun = mean, geom = "point", color = "black")           # Plot mean
```

