---
title: "Chlorophyll_a-c"
author: "EL Strand"
date: "7/17/2020"
output: html_document
---

Clears workspace. 
```{r}
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")
if (!require("dplyr")) install.packages("dplyr")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("gridExtra")) install.packages("gridExtra")

# load packages
library(tidyverse)
library(plotrix)
library(dplyr)
library(Rmisc)
library(ggplot2)
library(gridExtra)
```

Reading in datafiles and merging into one large dataframe. 
```{r}
df.platemaps <-
  list.files(path = 'Physiology_variables/Chlorophyll-a_c',pattern = "platemap.csv", full.names = TRUE) %>% 
  # list files in directory following a particular pattern
  set_names(.) %>% # get the column names
  map_dfr(read.csv, .id = "file.ID") %>% # join all files together in one data frame by file ID
 group_by(file.ID) 

df.platemaps$file.ID <- gsub("Physiology_variables/Chlorophyll-a_c/", "", df.platemaps$file.ID)
df.platemaps$file.ID <- gsub("_platemap.csv", "", df.platemaps$file.ID)

df.data <-
  list.files(path = 'Physiology_variables/Chlorophyll-a_c',pattern = "data.csv", full.names = TRUE) %>% 
  # list files in directory following a particular pattern
  set_names(.) %>% # get the column names
  map_dfr(read.csv, .id = "file.ID") %>% # join all files together in one data frame by file ID
 group_by(file.ID)

df.data$file.ID <- gsub("Physiology_variables/Chlorophyll-a_c/", "", df.data$file.ID)
df.data$file.ID <- gsub("_data.csv", "", df.data$file.ID)


# Merge platemap and data for each plate
df <- left_join(df.data,df.platemaps)
str(df)
```

# Calculate chlorophyll concentrations
```{r}
# average all technical replicates for each the sample duplicates per plate
df <- df %>%
  filter(!is.na(Plug_ID)) #remove empty wells (colony_id is NA)
  
df.1 <- df %>%
  dplyr::group_by(file.ID, Plug_ID) %>% #group by day and sample
  dplyr::summarise(mean_A630 = mean(X630), #calculate means
                   mean_A663 = mean(X663),
                   mean_A750 = mean(X750))
  

# get the acetone blank 750 absorbace for each file (i.e., plate), and subtract from 630 and 663 values for each sample
df.1 <- df.1 %>%
  dplyr::mutate(adj630 = mean_A630 - mean_A750,
         adj663 = mean_A663 - mean_A750)

# calculate chla and chlc2 values based on equations from Jeffrey and Humphrey 1975
# units µg/ml
df.1 <- df.1 %>%
  dplyr::mutate(chla.ug.ml = (11.43 * adj663) - (0.64 * adj630),
        chlc2.ug.ml = (27.09 * adj630) - (3.63 * adj663))

# path length adjustments for chlorophyll in 96 well quartz plate (0.584 cm)
df.1 <- df.1 %>%
  dplyr::mutate(chla.ug.ml = chla.ug.ml*0.584,
        chlc2.ug.ml = chlc2.ug.ml*0.584)
```


Standardize to surface area and homogenate volume. 
```{r}
# Load homogenate volume
# Load surface area
sa <- read_csv("Physiology_variables/Respirometry/coral_surface_area.csv") %>% dplyr::rename(Plug_ID = Sample_ID) %>%
  select(Plug_ID, surface.area.cm2)
sa$Plug_ID <- as.factor(sa$Plug_ID)
str(sa)

df.chl <- left_join(sa, df.1)

# Load homogenate volume
homog.vol <- read.csv("Physiology_variables/homogenate_vol.csv") %>%
  select(Plug_ID, vol_mL)
homog.vol$Plug_ID <- as.factor(homog.vol$Plug_ID)
str(homog.vol)

df.chl <- left_join(df.chl, homog.vol)

# Account for homogenate volume and normalize to surface area
df.chl <- df.chl %>%
  mutate(chla.ug.cm2 = (chla.ug.ml * vol_mL) / surface.area.cm2,
         chlc2.ug.cm2 = (chlc2.ug.ml * vol_mL) / surface.area.cm2)

# remove blanks and NAs
df.chl <- filter(df.chl, !Plug_ID %in% c("NA", "BK"))
```

Reading in master fragment information. 
```{r}
Master_fragment <- read.csv("Environmental_data/Master_Fragment.csv", header=T, sep=",", na.string="NA") #read in data file
Master_fragment$Plug_ID <- as.character(Master_fragment$Plug_ID)
Master_fragment <- Master_fragment %>% select(Plug_ID, Species, Timepoint, Temperature, Treatment, CO2)
str(df.chl)

Chl.plate.numbers <- read_csv("Physiology_variables/Chlorophyll-a_c/Chl_total_platemaps.csv") %>% 
  select(-Well) %>% distinct()

Chl_final <- left_join(df.chl, Master_fragment, by="Plug_ID")  
Chl_final$Timepoint <- factor(Chl_final$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week","12 week","16 week"))

Chl_final <- full_join(Chl.plate.numbers, Chl_final, by="Plug_ID") %>% na.omit(surface.area.cm2)

Chl_final %>% 
  select(Plug_ID, Species, Timepoint, Temperature, Treatment, CO2, chla.ug.cm2, chlc2.ug.cm2) %>%
  write_csv(path = "Output/chlorophyll.csv")
```


Checking for duplicates. 
```{r}
Chl.duplicates <- Chl_final %>% dplyr::group_by(Plug_ID) %>% 
  dplyr::summarise(n = n()) %>% filter(n > 1)
Chl.duplicates
# output should be zero bc NA and blanks taken out 
```

Checking for effect of plate 
```{r}
chlplotting <- Chl_final %>% na.omit()

ggplot(chlplotting, aes(x=Date, y=chlc2.ug.cm2, fill=Treatment)) + geom_boxplot(alpha=0.5) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + xlab("") + theme_bw() + 
  theme(legend.position = "right") + 
  facet_wrap(~Species, scales = "free_y") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(chlplotting, aes(x=Date, y=chlc2.ug.cm2, fill=Species)) + geom_boxplot(alpha=0.5) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + xlab("") + theme_bw() + 
  theme(legend.position = "right") + 
  facet_wrap(~Treatment, scales = "free_y") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



Calculating mean values for each timepoint and treatment group.
```{r}
mean.chl.a <- Chl_final %>%
  dplyr::group_by(Species, Timepoint, Temperature, CO2, Treatment) %>%
  dplyr::summarise(mean= mean(chla.ug.cm2), #mean 
                   N = sum(chla.ug.cm2), # sample size
                   se = sd(chla.ug.cm2)/sqrt(N)) #SE

mean.chl.c <- Chl_final %>%
  dplyr::group_by(Species, Timepoint, Temperature, CO2, Treatment) %>%
  dplyr::summarise(mean= mean(chlc2.ug.cm2), #mean 
                   N = sum(chlc2.ug.cm2), # sample size
                   se = sd(chlc2.ug.cm2)/sqrt(N)) #SE
```

Plotting timeseries
```{r}
cols <- c("lightblue", "blue", "pink", "red")

Fig.chla <- ggplot(mean.chl.a, aes(x=Timepoint, y=mean, group=Treatment)) + 
  geom_line(aes(colour=Treatment, group=Treatment), position = position_dodge(width = 0.1), alpha=1) + # colour, group both depend on cond2
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, size=1, position = position_dodge(width = 0.1)) +
  geom_point(aes(colour=Treatment), size = 3, position = position_dodge(width = 0.1)) +
  scale_colour_manual(values=cols) +
  xlab("Timepoint") +
  ylab(expression(Chl-a ~(µg ~cm^-2))) +
  #ylim(-0.5,1.5) +
  theme_bw() + #Set the background color
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), #Set the text angle
        axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        legend.position = "none", #remove legend
        plot.background=element_blank())+  #Set the plot background
  ggtitle("A) Chla") +
  facet_grid(rows=vars(Species), scales="free_y") +
  geom_vline(xintercept = c(7.2), colour="black", linetype="dotted")+
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 12, 
                                  hjust = 0))
Fig.chla

Fig.chlc2 <- ggplot(mean.chl.c, aes(x=Timepoint, y=mean, group=Treatment)) + 
  geom_line(aes(colour=Treatment, group=Treatment), position = position_dodge(width = 0.1), alpha=1) + # colour, group both depend on cond2
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="darkgray", width=0, size=1, position = position_dodge(width = 0.1)) +
  geom_point(aes(colour=Treatment), size = 3, position = position_dodge(width = 0.1)) +
  scale_colour_manual(values=cols) +
  xlab("Timepoint") +
  ylab(expression(Chl-c2 ~(µg ~cm^-2))) +
  #ylim(-0.5,1.5) +
  theme_bw() + #Set the background color
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), #Set the text angle
        axis.line = element_line(color = 'black'), #Set the axes color
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        legend.position = "none", #remove legend
        plot.background=element_blank())+  #Set the plot background
  ggtitle("B) Chlc2") +
  facet_grid(rows=vars(Species), scales="free_y") +
  geom_vline(xintercept = c(7.2), colour="black", linetype="dotted")+
  theme(plot.title = element_text(face = 'bold.italic', 
                                  size = 12, 
                                  hjust = 0))
Fig.chlc2

chl.Figs <- arrangeGrob(Fig.chla, Fig.chlc2, ncol=2)
ggsave(file="Output/Chlorophyll.pdf", chl.Figs, width = 11, height = 6, units = c("in"))
```

Correlation of chl and color score
```{r}


```


############## Bleaching Fragment Test prior to running the rest of the samples.

This was a test batch of 12 samples to make sure we would get a signal from both healthy and bleached samples. 

Reading in bleaching test information. 
```{r}
Phys_testing <- read.csv("Physiology_variables/phys_testing.csv")
Phys_testing$Plug_ID <- as.character(Phys_testing$Plug_ID)

Chl_testing <- full_join(Chl_final, Phys_testing, by="Plug_ID") %>% filter(!is.na(status))

Chl_testing <- gather(Chl_testing, "Measurement", "Value", 9:10) # changing columns to rows to allow for easier facet_grid in graphing section
```

Graphing bleach test data. 
```{r}
Bleached1.chl <- ggplot(Chl_testing, aes(x=status, y=Value, fill=status)) + geom_boxplot(alpha=0.5) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + xlab("") + theme_bw() + 
  theme(legend.position = "left") + 
  facet_grid(Measurement ~ Species, scales = "free")
Bleached1.chl

Bleached2.chl <- ggplot(Chl_testing, aes(x=bleaching_score, y=Value, color=status)) + geom_point() + 
  xlab("Bleaching Score") + theme_bw() + 
  facet_grid(Measurement ~ Species, scales = "free") + 
  theme(legend.position = "left")
Bleached2.chl

Chl.Blch.Figs <- arrangeGrob(Bleached1.chl, Bleached2.chl, ncol=2)

ggsave(file="Output/Phystest_Chl.pdf", Chl.Blch.Figs, width = 15, height = 10, units = c("in"))
```




