---
title: "Physiological_variability"
author: "EL Strand"
date: "7/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(data.table)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

Based on Tanner et al 2019.

Reading in growth and bleaching score dataset.
```{r}
## Bleaching Score
Bleaching_score <- read.csv("Physiology_variables/Bleaching_Score/Bleaching_Score.csv", header=T, sep=",", na.string="NA") 
Bleaching_score$Group <- as.character(Bleaching_score$Group)
Bleaching_score$PLUG.ID <- as.character(Bleaching_score$PLUG.ID)

Bleaching_variability <- reshape(Bleaching_score, timevar = "Timepoint",drop = c("Group", "SpGroup"), idvar=c("PLUG.ID","Species", "Temperature", "CO2", "Treatment"), direction="wide")

```

Transforming the dataframes into a matrix format of groups and response variable values.
```{r}
## Bleaching Score
Bleaching_score <- spread(Bleaching_score, Group, Bleaching.Score) 
Bleaching_variability <- select(Bleaching_score, contains("Week")) 

# coerce to data.table in place, move NAs to the bottom of each column, maintain the original order of non-NA values
Bleaching_variability <- data.table(Bleaching_variability)[, lapply(.SD, function(x) x[order(is.na(x))])]

# Bleaching_variability <- Bleaching_variability[!Bleaching_variability[, Reduce(`&`, lapply(.SD, is.na))]]
```


Calculating the Median Absolute Deviation (MAD) for each treatment group at each timepoint.
MAD is a robust estimate of variance calculated using the absolute deviations of each sampled value (Xi) from the median (∼X) of the respective treatment group (Hampel, 1974).
https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/mad
```{r}
## Creating a function for the MAD calculation
Median_abs_dev <- function(x) {
  MAD <- mad(x, center = median(x), constant = 1, na.rm = TRUE, low = FALSE, high = FALSE)
  return(MAD)
}

## Creating a loop to calculate MAD for every column in the Bleaching_variability dataframe
MAD <- vector("integer", ncol(Bleaching_variability)) # 1. defining output as a new list
for (i in seq_along(Bleaching_variability)) {            # 2. sequence: for every column in the dataframe, do what is defined in #3
  MAD[[i]] <- Median_abs_dev(Bleaching_variability[[i]])      # 3. body: function desired during the loop
}

MAD_values <- data.frame(MAD, colnames(Bleaching_variability)) %>% dplyr::rename(Group = colnames.Bleaching_variability.) %>%
  separate(Group, c("Timepoint", "Treatment", "Species"))


mad(Bleaching_variability$`Week1 ATAC Montipora`, center = median(Bleaching_variability$`Week1 ATAC Montipora`), constant = 1, na.rm = TRUE, low = FALSE, high = FALSE)
```
