---
title: "PI_Curves"
author: "EL Strand"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing and loading in required libraries. 
```{r}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
#if ("phytotools" %in% rownames(installed.packages()) == 'FALSE') install.packages('phytotools') 

#Read in required libraries
##### Include Versions of libraries
#install_github('colin-olito/LoLinR')
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
#library('phytotools')
```

Setting path and loading in required files. 
```{r}
path.p<-"Physiology_variables/PI_Curves/Raw_Data/" #the location of all your respirometry files 

#bring in the files
file.names <- list.files(path = path.p, pattern = "csv$") #list all csv file names in the folder
Photo.R <- data.frame(matrix(NA, nrow=length(file.names)*2, ncol=4)) #generate a 3 column dataframe with specific column names
colnames(Photo.R) <- c("Fragment.ID","Intercept", "µmol.L.sec", "Temp")

#add file names that include the subdirectory name
file.names.full <- list.files(path = path.p, pattern = "csv$", recursive = TRUE) #list all csv file names in the folder and subfolders

#Load Sample meta info Info
Sample.Info <- read.csv(file="Physiology_variables/PI_Curves/PI_Curve_Sample_Info_metadata.csv", header=T) #read sample.info data
Sample.Info$Fragment.ID <- paste(Sample.Info$Fragment.ID,"_",Sample.Info$Light_Level,sep = "")
```

Subset the data by light step using time breaks in the data. 
```{r}
for(i in 1:length(file.names.full)) { # for every file in list start at the first and run this following function
  Photo.Data1 <-read.table(file.path(path.p,file.names.full[i]),  header=T, sep=",", skip=1, na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Photo.Data1  <- as.data.frame(cbind(Photo.Data1$Time, Photo.Data1$Value, Photo.Data1$Temp)) #subset columns of interest
  colnames(Photo.Data1) <- c("Time", "Value", "Temp")
  Photo.Data1$Time <- as.POSIXct(Photo.Data1$Time,format="%H:%M") #convert time from character to time
  brk <- unique(as.POSIXct(Sample.Info$Start.time,format="%H:%M"))
  Step1 <- subset(Photo.Data1, Photo.Data1$Time > brk[1] & Photo.Data1$Time < brk[2])
  Step2 <- subset(Photo.Data1, Photo.Data1$Time > brk[2] & Photo.Data1$Time < brk[3])
  Step3 <- subset(Photo.Data1, Photo.Data1$Time > brk[3] & Photo.Data1$Time < brk[4])
  Step4 <- subset(Photo.Data1, Photo.Data1$Time > brk[4] & Photo.Data1$Time < brk[5])
  Step5 <- subset(Photo.Data1, Photo.Data1$Time > brk[5] & Photo.Data1$Time < brk[6])
  Step6 <- subset(Photo.Data1, Photo.Data1$Time > brk[6] & Photo.Data1$Time < brk[7])
  Step7 <- subset(Photo.Data1, Photo.Data1$Time > brk[7] & Photo.Data1$Time < brk[8])
  Step8 <- subset(Photo.Data1, Photo.Data1$Time > brk[8] & Photo.Data1$Time < brk[9])
  Step9 <- subset(Photo.Data1, Photo.Data1$Time > brk[9] & Photo.Data1$Time < brk[10])
  Step10 <- subset(Photo.Data1, Photo.Data1$Time > brk[10])
  lt.levs <- list(Step1, Step2, Step3, Step4, Step5, Step6, Step7, Step8, Step9, Step10) #list levels of segmentation
  
  for(j in 1:length(lt.levs)){    
    Photo.Data <- as.data.frame(lt.levs[j])
    n<-dim(Photo.Data)[1] #identify length of data
    Photo.Data <-Photo.Data [120:(n-3),] #start at data point ~2 minute in to avoid excess noise from start of run and remove last 3 lines containing text
    n<-dim(Photo.Data)[1] #list length of trimmed data
    Photo.Data$sec <- as.numeric(1:n) #set seconds by one from start to finish of run
    Photo.Data$Value <- as.numeric(as.character(Photo.Data$Value)) #save O2 data as numeric
    Photo.Data$Temp <- as.numeric(as.character(Photo.Data$Temp)) #save O2 data as numeric
    
    #Save plot prior to and after data thinning to make sure thinning is not too extreme
    rename <- sub("_.*", "", file.names[i])
    pdf(paste0("Physiology_variables/PI_curves/Output/Data_thinning/",rename,"_",j,"thinning.pdf"))
    par(omi=rep(0.3, 4)) #set size of the outer margins in inches
    par(mfrow=c(1,2)) #set number of rows and columns in multi plot graphic
    plot(Value ~ sec, data=Photo.Data , xlab='Time (seconds)', ylab=substitute(' O'[2]~' (µmol/L)'),  axes=FALSE) #plot data as a function of time
    usr  <-  par('usr')
    rect(usr[1], usr[3], usr[2], usr[4], col='grey90', border=NA)
    whiteGrid()
    box()
    points(Photo.Data$Value ~ Photo.Data $sec, pch=16, col=transparentColor('dodgerblue2', 0.6), cex=1.1)
    axis(1)
    axis(2, las=1)
    
    #set data reduction thinning parameter
    thin <- 40
    #save original unthinned data
    Photo.Data.orig<-Photo.Data
    
    Photo.Data   <-  thinData(Photo.Data.orig , by=thin)$newData1 #thin data 
    Photo.Data $sec <- as.numeric(rownames(Photo.Data)) #maintain numeric values for time
    Photo.Data$Temp <- NA # add a new column to fill with the thinned data
    Photo.Data$Temp <-  thinData(Photo.Data.orig, xy=c(2,3),by=thin)$newData1[,2] #thin data  for the temp values
    plot(Value ~ sec, data=Photo.Data , xlab='Time (seconds)', ylab=substitute(' O'[2]~' (µmol/L)'), type='n', axes=FALSE) #plot thinned data
    usr  <-  par('usr')
    rect(usr[1], usr[3], usr[2], usr[4], col='grey90', border=NA)
    whiteGrid()
    box()
    points(Photo.Data $Value ~ Photo.Data $sec, pch=16, col=transparentColor('dodgerblue2', 0.6), cex=1.1)
    axis(1)
    axis(2, las=1)
    dev.off()
    
    Regs  <-  rankLocReg(xall=Photo.Data$sec, yall=Photo.Data$Value, alpha=0.2, 
                         method="pc", verbose=TRUE) 
    pdf(paste0("Physiology_variables/PI_Curves/Output/Regressions/",rename,"_",j,"regression.pdf"))
    plot(Regs)
    dev.off()
    
    s <- seq(0,nrow(Photo.R),length(lt.levs)) #to order the file output sequence in correct order in data frame
    Photo.R[j+s[i],2:3] <- Regs$allRegs[1,c(4,5)] #inserts slope and intercept in the dataframe
    Photo.R[j+s[i],1] <- paste0(rename,"_",j) #stores the file name in the Date column
    Photo.R[j+s[i],4] <- mean(Photo.Data$Temp, na.rm=T)  #stores the Temperature in the Temp.C column
  }
}
```

Exporting rates and merging with sample info. 
```{r}
#view rates
Photo.R
write.csv(Photo.R,"Physiology_variables/PI_Curves/Output/PI_Curve_rates.csv")

#Merge rates with sample info
Data <- merge(Photo.R, Sample.Info, by="Fragment.ID")

#correct for the size of the chamber
Data$micromol.s <- Data$µmol.L.sec * Data$Chamber.Vol.L
```

Subtracting the blanks from each step.
```{r}
#calculate the average of the blanks from each time step
blnks <- subset(Data, Sample.Type=="Blank")
blnks <-mean.blnks <- aggregate(micromol.s ~ Light_Level, data=blnks, FUN=mean)
Data <- merge(Data, blnks, by="Light_Level")
colnames(Data)[colnames(Data) == 'micromol.s.x'] <- 'sample.micromol.s'
colnames(Data)[colnames(Data) == 'micromol.s.y'] <- 'blank.micromol.s'

#subtract the average of the blanks from each time step
Data$corr.micromol.s <- Data$sample.micromol.s - Data$blank.micromol.s
Data$micromol.cm2.s <- Data$corr.micromol.s/Data$Surf.Area.cm2
Data$micromol.cm2.h <- Data$micromol.cm2.s*3600
```

Export corrected rates. 
```{r}
PA <- subset(Data, Species=="Pacuta")
write.csv(PA,"Physiology_variables/PI_Curves/Output/PI_Curve_rates_Pacuta.csv")
MC <- subset(Data, Species=="Mcapitata")
write.csv(MC,"Physiology_variables/PI_Curves/Output/PI_Curve_rates_Mcapitata.csv")
```

Plotting PI curves. 
```{r}
pdf("Physiology_variables/PI_Curves/Output/BothSpecies_PICurves.pdf")
par(mfrow=c(2,1))

#Pocillopora acuta
PAR <- as.numeric(PA$Light_Value)
Pc <- as.numeric(PA$micromol.cm2.h)
plot(PAR,Pc,xlab="", ylab="", xlim=c(0,max(PAR)),cex.lab=0.8,cex.axis=0.8,cex=1, main="A) P. acuta", adj = 0.05) #set plot info
mtext(expression("Irradiance ("*mu*"mol photons "*m^-2*s^-1*")"),side=1,line=3.3,cex=1) #add labels
mtext(expression(Rate*" ("*mu*"mol "*O[2]*" "*cm^-2*h^-1*")"),side=2,line=2,cex=1) #add labels

PA.mod <- fitJP(PAR, Pc, normalize = FALSE, lowerlim = c(0, 1), upperlim = c(100, 1000), fitmethod=c("Nelder-Mead"))
A.PA <- PA.mod$alpha[1]
A.PA
Ik.PA <- PA.mod$ek[1]
Ik.PA

#Add model fit
E <- seq(-50,1500,by=1)
P <- PA.mod$alpha[1]*PA.mod$ek[1]*tanh(E/PA.mod$ek[1])
lines(E,P)
lines(cbind(PA.mod$ek[1], P),lwd=2,col="red") #add Ik line

# Montipora capitata
PAR <- as.numeric(MC$Light_Value)
Pc <- as.numeric(MC$micromol.cm2.h)
plot(PAR,Pc,xlab="", ylab="", xlim=c(0,max(PAR)),cex.lab=0.8,cex.axis=0.8,cex=1, main="B) M. capitata", adj = 0.05) #set plot info
mtext(expression("Irradiance ("*mu*"mol photons "*m^-2*s^-1*")"),side=1,line=3.3,cex=1) #add labels
mtext(expression(Rate*" ("*mu*"mol "*O[2]*" "*cm^-2*h^-1*")"),side=2,line=2,cex=1) #add labels

MC.mod <- fitJP(PAR, Pc, normalize = FALSE, lowerlim = c(0, 1), upperlim = c(100, 1000), fitmethod=c("Nelder-Mead"))
A.MC <- MC.mod$alpha[1]
A.MC
Ik.MC <- MC.mod$ek[1]
Ik.MC

#Add model fit
E <- seq(-50,1500,by=1)
P <- MC.mod$alpha[1]*MC.mod$ek[1]*tanh(E/MC.mod$ek[1])
lines(E,P, ,col="red")
lines(cbind(MC.mod$ek[1], P),lwd=2,col="red") #add Ik line

dev.off()
```

Nonlinear Least Squares regression of a non-rectangular hyperbola (Marshall & Biscoe, 1980)
```{r}
#Pocillopora acuta
PAR <- as.numeric(PA$Light_Value)
Pc <- as.numeric(PA$micromol.cm2.h)
plot(PAR,Pc,xlab="", ylab="", xlim=c(0,max(PAR)), ylim=c(-0.4, 1.0), cex.lab=0.8,cex.axis=0.8,cex=1, main="A) P. acuta", adj = 0.05) #set plot info
mtext(expression("Irradiance ("*mu*"mol photons "*m^-2*s^-1*")"),side=1,line=3.3,cex=1) #add labels
mtext(expression(Rate*" ("*mu*"mol "*O[2]*" "*cm^-2*h^-1*")"),side=2,line=2,cex=1) #add labels

#fit a model using a Nonlinear Least Squares regression of a non-rectangular hyperbola (Marshall & Biscoe, 1980)
curve.nlslrc.PA = nls(Pc ~ (1/(2*theta))*(AQY*PAR+Am-sqrt((AQY*PAR+Am)^2-4*AQY*theta*Am*PAR))-Rd,start=list(Am=(max(Pc)-min(Pc)),AQY=0.05,Rd=-min(Pc),theta=0.64)) 

my.fit.PA <- summary(curve.nlslrc.PA ) #summary of model fit

#draw the curve using the model fit
curve.fitting.PA <- curve((1/(2*summary(curve.nlslrc.PA)$coef[4,1]))*(summary(curve.nlslrc.PA)$coef[2,1]*x+summary(curve.nlslrc.PA)$coef[1,1]-sqrt((summary(curve.nlslrc.PA)$coef[2,1]*x+summary(curve.nlslrc.PA)$coef[1,1])^2-4*summary(curve.nlslrc.PA)$coef[2,1]*summary(curve.nlslrc.PA)$coef[4,1]*summary(curve.nlslrc.PA)$coef[1,1]*x))-summary(curve.nlslrc.PA)$coef[3,1],lwd=2,col="blue",add=T)
#Extract the parameters

#Amax (max gross photosytnthetic rate) 
Pmax.gross <- my.fit.PA$parameters[1]

#AQY (apparent quantum yield) alpha  
AQY <- my.fit.PA$parameters[2]

#Rd (dark respiration)
Rd <- my.fit.PA$parameters[3]

# Ik light saturation point
Ik <- Pmax.gross/AQY

# Ic light compensation point
Ic <- Rd/AQY

# Net photosynthetic rates
Pmax.net <- Pmax.gross - Rd

#output parameters into a table
PI.Output <- rbind(Pmax.gross, Pmax.net, -Rd, AQY,Ik,Ic)
row.names(PI.Output) <- c("Pg.max","Pn.max","Rdark","alpha", "Ik", "Ic")

# Montipora capitata
PAR <- as.numeric(MC$Light_Value)
Pc <- as.numeric(MC$micromol.cm2.h)
#plot(PAR,Pc,xlab="", ylab="", xlim=c(0,max(PAR)),ylim=c(-0.4,1.0),cex.lab=0.8,cex.axis=0.8,cex=1, main="B) M. capitata", adj = 0.05) #set plot info
#mtext(expression("Irradiance ("*mu*"mol photons "*m^-2*s^-1*")"),side=1,line=3.3,cex=1) #add labels
#mtext(expression(Rate*" ("*mu*"mol "*O[2]*" "*cm^-2*h^-1*")"),side=2,line=2,cex=1) #add labels

#fit a model using a Nonlinear Least Squares regression of a non-rectangular hyperbola (Marshall & Biscoe, 1980)
curve.nlslrc.MC = nls(Pc ~ (1/(2*theta))*(AQY*PAR+Am-sqrt((AQY*PAR+Am)^2-4*AQY*theta*Am*PAR))-Rd,start=list(Am=(max(Pc)-min(Pc)),AQY=0.05,Rd=-min(Pc),theta=0.6)) 

my.fit.MC <- summary(curve.nlslrc.MC) #summary of model fit

#draw the curve using the model fit
#curve.fitting.MC <- curve((1/(2*summary(curve.nlslrc.MC)$coef[4,1]))*(summary(curve.nlslrc.MC)$coef[2,1]*x+summary(curve.nlslrc.MC)$coef[1,1]-sqrt((summary(curve.nlslrc.MC)$coef[2,1]*x+summary(curve.nlslrc.MC)$coef[1,1])^2-4*summary(curve.nlslrc.MC)$coef[2,1]*summary(curve.nlslrc.MC)$coef[4,1]*summary(curve.nlslrc.MC)$coef[1,1]*x))-summary(curve.nlslrc.MC)$coef[3,1],lwd=2,col="blue",add=T)
#lines(cbind(Pmax.gross/AQY, P),lwd=2,col="gray", lty=3) #add Ik line
#Extract the parameters

#Amax (max gross photosytnthetic rate) 
Pmax.gross <- my.fit.MC$parameters[1]

#AQY (apparent quantum yield) alpha  
AQY <- my.fit.MC$parameters[2]

#Rd (dark respiration)
Rd <- my.fit.MC$parameters[3]

# Ik light saturation point
Ik <- Pmax.gross/AQY

# Ic light compensation point
Ic <- Rd/AQY

# Net photosynthetic rates
Pmax.net <- Pmax.gross - Rd

#output parameters into a table
PI.Output <- rbind(Pmax.gross, Pmax.net, -Rd, AQY,Ik,Ic)
row.names(PI.Output) <- c("Pg.max","Pn.max","Rdark","alpha", "Ik", "Ic")
PI.Output
```

Plot input data and model fit
```{r}
plot(PAR,Pc,xlab="", ylab="", xlim=c(0,max(PAR)), ylim=c(-0.4,1.0),cex.lab=0.8,cex.axis=0.8,cex=1,main="B) M. capitata", adj = 0.05) #set plot info
mtext(expression("Irradiance ("*mu*"mol photons "*m^-2*s^-1*")"),side=1,line=3.3,cex=1) #add labels
mtext(expression(Rate*" ("*mu*"mol "*O[2]*" "*cm^-2*h^-1*")"),side=2,line=2,cex=1) #add labels
curve((1/(2*summary(curve.nlslrc.MC)$coef[4,1]))*(summary(curve.nlslrc.MC)$coef[2,1]*x+summary(curve.nlslrc.MC)$coef[1,1]-sqrt((summary(curve.nlslrc.MC)$coef[2,1]*x+summary(curve.nlslrc.MC)$coef[1,1])^2-4*summary(curve.nlslrc.MC)$coef[2,1]*summary(curve.nlslrc.MC)$coef[4,1]*summary(curve.nlslrc.MC)$coef[1,1]*x))-summary(curve.nlslrc.MC)$coef[3,1],lwd=2,col="blue",add=T)
#add fit line
#lines(cbind(curve.fitting$x, PI.Output[2]),lwd=2,col="yellow") #add Pnet max line
abline(v=Ik,lwd=2,col="gray", lty=3)
#lines(cbind(PI.Output[5], curve.fitting.MC$y),lwd=2,col="gray", lty=3) #add Ik line
#abline(PI.Output[3], PI.Output[4],lwd=2,col="green") #add alpha line
#legend( max(PAR)-300, min(Pc)+5, legend=c("Curve Fit", "Pnet max", "Ik", "alpha"), col=c("blue", "yellow", "red", "green"), lty=1, cex=1, box.lty = 0 ) #add a legend
dev.off()

PI.Output.MC<- as.data.frame(PI.Output)
colnames(PI.Output.MC) <- c("Mcapitata")

PI.Output.Values <- cbind(PI.Output.PA, PI.Output.MC)
PI.Output.Values
write.csv(PI.Output.Values,"Physiology_variables/PI_Curves/Output/PI_Curve_Output.csv")
```


E5 model: nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=(max(.$Pc)-min(.$Pc)),  AQY=0.001, Rd=-min(.$Pc))). 

nls(Pc ~ (1/(2*theta))*(AQY*PAR+Am-sqrt((AQY*PAR+Am)^2-4*AQY*theta*Am*PAR))-Rd,start=list(Am=(max(Pc)-min(Pc)),AQY=0.05,Rd=-min(Pc),theta=0.6)) 



