---
title: "Soluble_Protein"
author: "EL Strand"
date: "8/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
```

Load in required libraries. 
```{r}
## install packages if you dont already have them
if (!require("dplyr")) install.packages("dplyr")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")

# load packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(broom)
```

Based on Pierce BCA Rapid Gold Protein Kit.

Reading in datafiles. 
```{r}
## Bleaching Test Plate 1 
plate1_platemap <- read.csv("Physiology_variables/Soluble_Protein/20200804_Plate1_SP_platemap.csv") 
  plate1_platemap$Plug_ID <- as.character(plate1_platemap$Plug_ID)
plate1_data <- read.csv("Physiology_variables/Soluble_Protein/20200804_plate1_SP.csv", header=T) 
  plate1_data$X480 <- gsub("OVRFLW", "NA", plate1_data$X480) # replacing OVRFLW valuew with NA
  plate1_data$X480 <- as.numeric(plate1_data$X480)

plate1 <- full_join(plate1_platemap, plate1_data, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID)) # joining two above dataframes; renaming X480 column; removing all wells with NA values

# new dataframe (standards) with the standard values and renaming that column
standards1 <- plate1 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID)
```

```{r}
## 20201007 Plate numbers 1, 2, 3 
Oct7_plate1 <- read.csv("Physiology_variables/Soluble_Protein/20201007_plate1_SP.csv")
  Oct7_plate1$X480 <- as.numeric(Oct7_plate1$X480)
  Oct7_plate1$plate.number <- rep(1,nrow(Oct7_plate1))
Oct7_platemap1 <- read.csv("Physiology_variables/Soluble_Protein/20201007_platemap1.csv")
  Oct7_platemap1$Plug_ID <- as.character(Oct7_platemap1$Plug_ID)

Oct7_plate1 <- full_join(Oct7_platemap1, Oct7_plate1, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID))

Oct7_plate2 <- read.csv("Physiology_variables/Soluble_Protein/20201007_plate2_SP.csv")
  Oct7_plate2$X480 <- as.numeric(Oct7_plate2$X480)
  Oct7_plate2$plate.number <- rep(2,nrow(Oct7_plate2))
Oct7_platemap2 <- read.csv("Physiology_variables/Soluble_Protein/20201007_platemap2.csv")
  Oct7_platemap2$Plug_ID <- as.character(Oct7_platemap2$Plug_ID)

Oct7_plate2 <- full_join(Oct7_platemap2, Oct7_plate2, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID))

Oct7_plate3 <- read.csv("Physiology_variables/Soluble_Protein/20201007_plate3_SP.csv")
  Oct7_plate3$X480 <- as.numeric(Oct7_plate3$X480)
  Oct7_plate3$plate.number <- rep(3,nrow(Oct7_plate3))
Oct7_platemap3 <- read.csv("Physiology_variables/Soluble_Protein/20201007_platemap3.csv")
  Oct7_platemap3$Plug_ID <- as.character(Oct7_platemap3$Plug_ID)

Oct7_plate3 <- full_join(Oct7_platemap3, Oct7_plate3, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID))

Oct7 <- union(Oct7_plate1, Oct7_plate2) %>% union(Oct7_plate3)
  Oct7.standards <- Oct7 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID) 
```


```{r}
## 20201014 Plate numbers 4, 5, 6
Oct14_plate1 <- read.csv("Physiology_variables/Soluble_Protein/20201014_plate1_SP.csv")
  Oct14_plate1$X480 <- as.numeric(Oct14_plate1$X480)
  Oct14_plate1$plate.number <- rep(4,nrow(Oct14_plate1))
Oct14_platemap1 <- read.csv("Physiology_variables/Soluble_Protein/20201014_platemap1.csv")
  Oct14_platemap1$Plug_ID <- as.character(Oct14_platemap1$Plug_ID)
Oct14_plate1 <- full_join(Oct14_platemap1, Oct14_plate1, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID))

Oct14_plate2 <- read.csv("Physiology_variables/Soluble_Protein/20201014_plate2_SP.csv")
  Oct14_plate2$X480 <- as.numeric(Oct14_plate2$X480)
  Oct14_plate2$plate.number <- rep(5,nrow(Oct14_plate2))
Oct14_platemap2 <- read.csv("Physiology_variables/Soluble_Protein/20201007_platemap2.csv")
  Oct14_platemap2$Plug_ID <- as.character(Oct14_platemap2$Plug_ID)
Oct14_plate2 <- full_join(Oct14_platemap2, Oct14_plate2, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID))

Oct14_plate3 <- read.csv("Physiology_variables/Soluble_Protein/20201014_plate3_SP.csv")
  Oct14_plate3$X480 <- as.numeric(Oct14_plate3$X480)
  Oct14_plate3$plate.number <- rep(6,nrow(Oct14_plate3))
Oct14_platemap3 <- read.csv("Physiology_variables/Soluble_Protein/20201014_platemap3.csv")
  Oct14_platemap3$Plug_ID <- as.character(Oct14_platemap3$Plug_ID)
Oct14_plate3 <- full_join(Oct14_platemap3, Oct14_plate3, by="Well") %>% dplyr::rename(Abs480 = X480) %>% filter(!is.na(Plug_ID))

Oct14 <- union(Oct14_plate1, Oct14_plate2) %>% union(Oct14_plate3)
  Oct14.standards <- Oct14 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID) 
```

Plot standard curves. 
```{r}
st.name <- c("StandardA", "StandardB", "StandardC", "StandardD", "StandardE", "StandardF", "StandardG", "StandardH", "StandardI")
concentration <- c("2000", "1500", "1000", "750", "500", "250", "125", "25", "0")

## Bleaching Test Plate 1 
standards1 <- standards1 %>%
  dplyr::group_by(Standard) %>%
  summarise(Abs480 = mean(Abs480)) %>%
  mutate(Abs480.adj = Abs480 - Abs480[Standard == "StandardI"])

standards1$BSA_ug.mL <- concentration[match(standards1$Standard, st.name)]
  standards1$BSA_ug.mL <- as.numeric(standards1$BSA_ug.mL)

SP_mod_lin1 <- lm(BSA_ug.mL ~ Abs480.adj, data = standards1)
summary(SP_mod_lin1)

std_curve_plot1 <- standards1 %>%
  ggplot(aes(x = BSA_ug.mL, y = Abs480.adj)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  geom_smooth(method = "lm") + 
  labs(title = paste("Adj R2 = ", signif(summary(SP_mod_lin1)$adj.r.squared, 5),
                     "; Soluble Protein Standard Curve Bleaching Plate 1"))
std_curve_plot1

## 20201007  
Oct7.standards <- Oct7.standards %>%
  dplyr::group_by(Standard) %>%
  summarise(Abs480 = mean(Abs480)) %>%
  mutate(Abs480.adj = Abs480 - Abs480[Standard == "StandardI"])
Oct7.standards$BSA_ug.mL <- concentration[match(Oct7.standards$Standard, st.name)]
  Oct7.standards$BSA_ug.mL <- as.numeric(Oct7.standards$BSA_ug.mL)
SP_mod_lin_Oct7 <- lm(BSA_ug.mL ~ Abs480.adj, data = Oct7.standards)
summary(SP_mod_lin_Oct7)
std_curve_plot2 <- Oct7.standards %>%
  ggplot(aes(x = Abs480.adj, y = BSA_ug.mL)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  geom_smooth(method = "lm") + 
  labs(title = paste("Adj R2 = ", signif(summary(SP_mod_lin_Oct7)$adj.r.squared, 5),
                     "; Soluble Protein Standard Curve Oct7 3 Plates"))
std_curve_plot2

## 20201014  
Oct14.standards <- Oct14.standards %>%
  dplyr::group_by(Standard) %>%
  summarise(Abs480 = mean(Abs480)) %>%
  mutate(Abs480.adj = Abs480 - Abs480[Standard == "StandardI"])
Oct14.standards$BSA_ug.mL <- concentration[match(Oct14.standards$Standard, st.name)]
  Oct14.standards$BSA_ug.mL <- as.numeric(Oct14.standards$BSA_ug.mL)
SP_mod_lin_Oct14 <- lm(BSA_ug.mL ~ Abs480.adj, data = Oct14.standards)
summary(SP_mod_lin_Oct14)
std_curve_plot3 <- Oct14.standards %>%
  ggplot(aes(x = BSA_ug.mL, y = Abs480.adj)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  geom_smooth(method = "lm") + 
  labs(title = paste("Adj R2 = ", signif(summary(SP_mod_lin_Oct14)$adj.r.squared, 5),
                     "; Soluble Protein Standard Curve Oct14 3 Plates"))
std_curve_plot3
```

The standard curve for the bleaching test sample has a high R squared value but the absorbance values look off. 

Calculating protein concentrations based on standard curve. 
```{r}
## Bleaching Test Plate 1
protein1 <- plate1 %>% filter(!grepl("Standard", Plug_ID)) %>%
  filter(!is.na(Plug_ID)) %>%
  group_by(Plug_ID) %>%
  summarise(Abs480 = mean(Abs480)) %>%
  mutate(Abs480.adj = Abs480 - standards1$Abs480[standards1$Standard == "StandardI"],  # Subtract blank standard (Standard I) from mean absorbance values
         prot_ug.mL = map_dbl(Abs480.adj, ~ predict(SP_mod_lin1, newdata = data.frame(Abs480.adj = .)))) # Use standard curve to convert absorbance to protein

std_curve_plot1 + 
  geom_point(data = protein1, aes(x = prot_ug.mL, y = Abs480.adj), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve 1 - bleaching test")

## 20201007 3 Plates
Oct7protein <- Oct7 %>% 
  filter(!is.na(Plug_ID)) %>%
  group_by(Plug_ID) %>%
  summarise(Abs480 = mean(Abs480)) %>% 
  mutate(Abs480.adj = Abs480 - Oct7.standards$Abs480[Oct7.standards$Standard == "StandardI"],  # Subtract blank standard (Standard I) from mean absorbance values
         prot_ug.mL = map_dbl(Abs480.adj, ~ predict(SP_mod_lin_Oct7, newdata = data.frame(Abs480.adj = .)))) # Use standard curve to convert absorbance to protein

std_curve_plot2 + 
  geom_point(data = Oct7protein, aes(x = Abs480.adj, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve 2 - Oct 7 3 Plates")

## 20201014 3 Plates
Oct14protein <- Oct14 %>% filter(!grepl("Standard", Plug_ID)) %>%
  filter(!is.na(Plug_ID)) %>%
  group_by(Plug_ID) %>%
  summarise(Abs480 = mean(Abs480)) %>% 
  mutate(Abs480.adj = Abs480 - Oct14.standards$Abs480[Oct14.standards$Standard == "StandardI"],  # Subtract blank standard (Standard I) from mean absorbance values
         prot_ug.mL = map_dbl(Abs480.adj, ~ predict(SP_mod_lin_Oct14, newdata = data.frame(Abs480.adj = .)))) # Use standard curve to convert absorbance to protein

std_curve_plot3 + 
  geom_point(data = Oct14protein, aes(x = prot_ug.mL, y = Abs480.adj), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve 3 - Oct 14 3 Plates")
```

Normalize to surface area. 
```{r}
sa <- read_csv("Physiology_variables/Respirometry/coral_surface_area.csv") %>% dplyr::rename(Plug_ID = Sample_ID) %>% select(Plug_ID, surface.area.cm2)
  sa$Plug_ID <- as.character(sa$Plug_ID)
Homogenate_volume <- read.csv("Physiology_variables/homogenate_vol.csv") %>% select(Plug_ID, vol_mL)
  Homogenate_volume$Plug_ID <- as.character(Homogenate_volume$Plug_ID)

Master_fragment <- read.csv("Environmental_data/Master_Fragment.csv", header=T, sep=",", na.string="NA") %>% select(Plug_ID, Species, Timepoint, Temperature, Treatment, CO2)
  Master_fragment$Plug_ID <- as.character(Master_fragment$Plug_ID)
  
metadata <- full_join(sa, Homogenate_volume, by='Plug_ID', type='left') %>% 
  full_join(Master_fragment)

## Join all protein plates together
protein <- Oct7protein

# change below to full join 
protein <- left_join(protein, metadata) %>%
  mutate(prot_ug = prot_ug.mL * vol_mL,
         prot_ug.cm2 = prot_ug / surface.area.cm2,
         prot_mg.cm2 = prot_ug.cm2/1000)


boxplot(prot_mg.cm2 ~Treatment*Species, data=protein, las=2)

```

# Write data to output file
```{r}
# Write protein data to output file
protein %>%
  select(Plug_ID, prot_ug.mL, prot_ug, prot_ug.cm2) %>%
  write_csv(., path = "Output/soluble_protein.csv")
```

# Exclude all values above 2000 ug.mL and below 0 ug.mL
These corals will have to be re-done and if still above the curve then dilute the sample because they are outside the standard curve. 

```{r}
dilute.corals <- filter(protein, prot_ug.mL > 2000)
re.do.corals <- filter(protein, prot_ug.mL < 0)
```



############# Bleaching Test Information prior to processing all of the samples ############# 

Reading in bleaching test info. 
```{r}
Phys_testing <- read.csv("Physiology_variables/phys_testing.csv")
  Phys_testing$Plug_ID <- as.character(Phys_testing$Plug_ID)
  
Protein_testing <- full_join(protein1, Phys_testing, by="Plug_ID")
```

Graphing bleaching test information.
```{r}
Bleached1 <- ggplot(Protein_testing, aes(x=status, y=prot_ug.cm2, fill=status)) + geom_boxplot(alpha=0.5) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + xlab("") + theme_bw() + 
  theme(legend.position = "") + 
  facet_wrap(~Species)
Bleached1

Bleached2 <- ggplot(Protein_testing, aes(x=bleaching_score, y=prot_ug.cm2)) + geom_point() + 
  xlab("Bleaching Score") + theme_bw() + 
  facet_wrap(~Species) + 
  theme(legend.position = "")
Bleached2
```


