---
title: "Statistics.Rmd"
author: "EL Strand"
date: "2/11/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clears workspace. 
```{r}
rm(list=ls())
```

Loading libraries. 
```{r}
library(ggplot2)
library(tidyverse)
library(plyr)
library(dplyr)
library(mgcv) # GAM function
library(ggfortify)
library(devtools)
install_github("vqv/ggbiplot")
library(lme4)
library(gridExtra)
library(vegan)
library(effects)
library(Rmisc)
library(lmerTest)
library(lessR)
library(car)
library(cowplot)
library(corrplot)
library(factoextra)
```

### Read in data files to analyze. 

```{r}
Master <- read.csv("Environmental_data/Master_Fragment.csv") %>% select(Plug_ID, Species, ANALYSIS, Timepoint, Treatment, Tank, Temperature, CO2) %>%
  filter(ANALYSIS == "Physiology" | ANALYSIS == "Physiology/Molecular")
  
Photo.Resp <- read.csv("Physiology_variables/Respirometry/Output/Resp_AllRates.csv") %>% select(Fragment.ID.x, Pnet_umol.cm2.hr, Rdark_umol.cm2.hr, Pgross_umol.cm2.hr) %>%
  dplyr::rename(Plug_ID = Fragment.ID.x)
  Photo.Resp$Plug_ID <- gsub("P-", "", Photo.Resp$Plug_ID)
  Photo.Resp$Plug_ID <- gsub("M-", "", Photo.Resp$Plug_ID)
  Photo.Resp <- Photo.Resp %>% separate(Plug_ID, c("Plug_ID", "Date")) %>% select(-Date)

Chlorophyll <-  read.csv("Output/chlorophyll.csv") %>% select(Plug_ID, chla.ug.cm2, chlc2.ug.cm2)

Biomass <- read.csv("Output/Tissue_Biomass.csv") %>% select(Plug_ID, partner, AFDW.mg.cm2)

Protein <- read.csv("Output/soluble_protein.csv") %>% select(Plug_ID, prot_mg.cm2)

Antioxidant <- read.csv("Output/antioxidant.csv") %>% select(Plug_ID, cre.umol.mgprot)

Color.Score <- read.csv("Physiology_variables/Bleaching_Score_Output.csv") %>% select(PLUG.ID, Bleaching.Score, Timepoint) %>% dplyr::rename(Plug_ID = PLUG.ID) %>% 
  dplyr::rename(Blch.Time = Timepoint)
Color.Score$Blch.Time <- gsub("^.{0,4}", "", Color.Score$Blch.Time) # taking out the "Week" in every row
Color.Score$Blch.Time <- paste(Color.Score$Blch.Time, "week") # adding "week" after each number in every row 

Growth <- read.csv("Physiology_variables/Growth_Buoyant_Weight/Growth_Rates.csv") %>% select(PLUG.ID, T0.g.d.cm2, T1.g.d.cm2, T2.g.d.cm2, T3.g.d.cm2, T4.g.d.cm2, T5.g.d.cm2,
                                                                                             T6.g.d.cm2, T7.g.d.cm2, T8.g.d.cm2) %>% 
  gather("Growth.Time", "Growth.Rate", 2:10) %>% na.omit(Growth.Rate) %>% dplyr::rename(Plug_ID=PLUG.ID) %>%
  mutate(Growth.Time = case_when(
    Growth.Time == "T0.g.d.cm2" ~ "1 week",
    Growth.Time == "T1.g.d.cm2" ~ "2 week", 
    Growth.Time == "T2.g.d.cm2" ~ "4 week",
    Growth.Time == "T3.g.d.cm2" ~ "6 week",
    Growth.Time == "T4.g.d.cm2" ~ "8 week",
    Growth.Time == "T5.g.d.cm2" ~ "10 week",
    Growth.Time == "T6.g.d.cm2" ~ "12 week",
    Growth.Time == "T7.g.d.cm2" ~ "14 week",
    Growth.Time == "T8.g.d.cm2" ~ "16 week")
  )
Growth$Plug_ID <- as.integer(Growth$Plug_ID)

# Cell.density

# come back to adding Color.Score and Growth back in and subsetting for where all three timepoint columns match 
# joining all datasets together, reshaping the AFDW partner into columns, renaming those columns, removing an NA column, and calculating S:H AFDW ratio
data <- join_all(list(Master, Photo.Resp, Chlorophyll, Biomass, Protein, Antioxidant), by='Plug_ID', type='left') %>%
  spread(partner, AFDW.mg.cm2) %>% dplyr::rename(Host_AFDW.mg.cm2 = Host) %>% dplyr::rename(Sym_AFDW.mg.cm2 = Sym) %>% select(-"<NA>") %>%
  mutate(Ratio_AFDW.mg.cm2=Sym_AFDW.mg.cm2/Host_AFDW.mg.cm2) 
data$Timepoint <- factor(data$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week", "12 week", "16 week"))

data <- data %>% 
  mutate(Phase = case_when(
    Timepoint == "Day 1" ~ "Acute",
    Timepoint == "Day 2" ~ "Acute",
    Timepoint == "1 week" ~ "Acute",
    Timepoint == "2 week" ~ "Chronic",
    Timepoint == "4 week" ~ "Chronic",
    Timepoint == "6 week" ~ "Chronic",
    Timepoint == "8 week" ~ "Chronic",
    Timepoint == "12 week" ~ "Recovery",
    Timepoint == "16 week" ~ "Recovery")) %>%
  unite(Group, Phase, Treatment, remove = FALSE, sep = " ") %>%
  write_csv(., file = "Output/results_physiology.csv")

timeseries.data <- data %>% select(-Pnet_umol.cm2.hr, -Rdark_umol.cm2.hr, -Pgross_umol.cm2.hr)

pacuta_full_data <- data %>% subset(Species == "Pacuta")
mcap_full_data <- data %>% subset(Species == "Mcapitata")

color.data <- full_join(Master, Color.Score, by = "Plug_ID") %>% na.omit() %>%
  write_csv(., file = "Output/results_colorscore.csv")
Growth.data <- full_join(Master, Growth, by = "Plug_ID") %>% na.omit() %>%
  write_csv(., file = "Output/results_growth.csv")

data <- data %>% filter(Ratio_AFDW.mg.cm2<3)
Ratio.means <- data %>% select(Plug_ID, Species, Timepoint, Treatment, Temperature, CO2, Ratio_AFDW.mg.cm2) %>% na.omit() %>% 
  summarySE(measurevar="Ratio_AFDW.mg.cm2", groupvars=c("Species", "Timepoint", "Treatment", "Temperature", "CO2")) %>% filter(N>2) %>%
  mutate(Species = case_when(
    Species == "Mcapitata" ~ "Montipora capitata",
    Species == "Pacuta" ~ "Pocillopora acuta"
  )) 
```

Plot AFDW Ratio. 
Taking out 1116 because it was an outlier. 

```{r}
cols <- c("lightblue", "blue", "salmon", "red3")

boxplot(data$Ratio_AFDW.mg.cm2)
filter(data, Ratio_AFDW.mg.cm2>3)
  
AFDW_Ratio_plot <- ggplot(Ratio.means, aes(x=Timepoint, y = Ratio_AFDW.mg.cm2, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(~ Species, scales = "free") +
  geom_errorbar(aes(ymin=Ratio_AFDW.mg.cm2-se, ymax=Ratio_AFDW.mg.cm2+se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Timepoint") + 
  theme_classic() + 
  #ylab("expression(AFDW ~(mg ~cm^-2))") +
  ylab("Ash Free Dry Weight S:H Ratio") +
  theme(legend.position = "left") +
  geom_vline(xintercept = c(7.2), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) +
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  strip.text.x = element_text(size = 12, color = "black", face = "bold.italic"),
  # Change axis line
  axis.line = element_line(colour = "black")); AFDW_Ratio_plot

ggsave(file="Output/Final_Figures/Tissue_Biomass-Ratio.pdf", AFDW_Ratio_plot, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Tissue_Biomass-Ratio.png", AFDW_Ratio_plot, width = 11, height = 6, units = c("in"))
```


# Univariate analysis

## Statistical Modeling
Fixed factors: Temperature, CO2, Species, Timepoint. 
Random factors: Tank and Plug ID

When adding GAM, use model.sel()  
MuMIn package

### Photosynthetic Rates and Respiration 

Raw data distribution view.
```{r}
hist(pacuta_full_data$Pnet_umol.cm2.hr)
hist(pacuta_full_data$Pgross_umol.cm2.hr)
hist(pacuta_full_data$Rdark_umol.cm2.hr)

hist(mcap_full_data$Pnet_umol.cm2.hr)
hist(mcap_full_data$Pgross_umol.cm2.hr)
hist(mcap_full_data$Rdark_umol.cm2.hr)
# mcap doesn't look normal, the models should account for this but might need to transform post models 
```

*P. acuta* linear mixed effect models.  
Net photosynthetic rates. 
```{r}
## Pnet
Pacuta_Pnet_LMER <- lmer(Pnet_umol.cm2.hr ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Pacuta_Pnet_LMER) in console to view diagnostic plots
summary(Pacuta_Pnet_LMER)
qqPlot(residuals(Pacuta_Pnet_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_Pnet_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
#leveneTest(residuals(Pacuta_Pnet_LMER))
leveneTest(pacuta_full_data$Pnet_umol.cm2.hr ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)
#bartlett.test(Pacuta_Pnet_LMER)


## significance of model
anova(Pacuta_Pnet_LMER, ddf="lme4", type='III')
Anova(Pacuta_Pnet_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Pnet_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

Gross photosynthetic rates. 
```{r}
## Pgross 
Pacuta_Pgross_LMER <- lmer(Pgross_umol.cm2.hr ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

## summary of the model chosen from above
## plot(Pacuta_Pgross_LMER) in console to view diagnostic plots
summary(Pacuta_Pgross_LMER)
qqPlot(residuals(Pacuta_Pgross_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_Pgross_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$Pgross_umol.cm2.hr ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_Pgross_LMER, ddf="lme4", type='III')
Anova(Pacuta_Pgross_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Pgross_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

```{r}
## Dark Resp
Pacuta_Resp_LMER <- lmer(Rdark_umol.cm2.hr ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Pacuta_Resp_LMER) in console to view diagnostic plots
summary(Pacuta_Resp_LMER)
qqPlot(residuals(Pacuta_Resp_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_Resp_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$Rdark_umol.cm2.hr ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_Resp_LMER, ddf="lme4", type='III')
Anova(Pacuta_Resp_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Resp_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

*M.capitata* linear mixed effect models 

Net photosynthetic rates. 
```{r}
## Pnet
Mcap_Pnet_LMER <- lmer(log(Pnet_umol.cm2.hr) ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Mcap_Pnet_LMER) in console to view diagnostic plots
summary(Mcap_Pnet_LMER)
qqPlot(residuals(Mcap_Pnet_LMER)) # qqplot and histogram 
hist(residuals(Mcap_Pnet_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$Pnet_umol.cm2.hr ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_Pnet_LMER, ddf="lme4", type='III')
Anova(Mcap_Pnet_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Pnet_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

```{r}
## Pgross
Mcap_Pgross_LMER <- lmer(log(Pgross_umol.cm2.hr) ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Mcap_Pnet_LMER) in console to view diagnostic plots
summary(Mcap_Pgross_LMER)
qqPlot(residuals(Mcap_Pgross_LMER)) # qqplot and histogram 
hist(residuals(Mcap_Pgross_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$Pgross_umol.cm2.hr ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_Pgross_LMER, ddf="lme4", type='III')
Anova(Mcap_Pgross_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Pgross_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

```{r}
## Dark Resp 
mcap_full_data$Rdarkplus1 <- mcap_full_data$Rdark_umol.cm2.hr+1
  
Mcap_Resp_LMER <- lmer(log(Rdarkplus1) ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Mcap_Resp_LMER) in console to view diagnostic plots
summary(Mcap_Resp_LMER)
qqPlot(residuals(Mcap_Resp_LMER)) # qqplot and histogram 
hist(residuals(Mcap_Resp_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$Rdark_umol.cm2.hr ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_Resp_LMER, ddf="lme4", type='III')
Anova(Mcap_Resp_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Resp_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```


### Chlorophyll Concentration

Raw data distribution view.
```{r}
hist(pacuta_full_data$chla.ug.cm2)
hist(mcap_full_data$chla.ug.cm2)
# mcap doesn't look normal, the models should account for this but might need to transform post models 
```

Model comparisons. 
*P. acuta* 

The full P.acuta dataset received the error `fixed-effect model matrix is rank deficient so dropping 3 columns / coefficients`. 

```{r}
## Subset out 12 and 16 week timepoints 
#Pacuta_stress_data <- pacuta_full_data %>% filter(Timepoint !="12 week" & Timepoint !="16 week")
  
Pacuta_Chl_LMER <- lmer(chla.ug.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Pacuta_Chl_LMER) in console to view diagnostic plots
summary(Pacuta_Chl_LMER)
qqPlot(residuals(Pacuta_Chl_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_Chl_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$chla.ug.cm2 ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_Chl_LMER, ddf="lme4", type='III')
Anova(Pacuta_Chl_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Chl_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

*M. capitata* 
```{r}
Mcap_Chl_LMER <- lmer(chla.ug.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values 

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Mcap_Chl_LMER)
qqPlot(residuals(Mcap_Chl_LMER)) # qqplot and histogram 
hist(residuals(Mcap_Chl_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$chla.ug.cm2 ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_Chl_LMER, ddf="lme4", type='III')
Anova(Mcap_Chl_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Chl_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

### Tissue Biomass 
Raw data distribution view.
```{r}
hist(pacuta_full_data$Host_AFDW.mg.cm2) # normal
hist(pacuta_full_data$Sym_AFDW.mg.cm2) # right/positively skewed 
hist(pacuta_full_data$Ratio_AFDW.mg.cm2) # extremely right/positively skewed

hist(mcap_full_data$Host_AFDW.mg.cm2) # right/positively skewed
hist(mcap_full_data$Sym_AFDW.mg.cm2) # right/positively skewed
hist(mcap_full_data$Ratio_AFDW.mg.cm2) # normal 
```

*P.acuta* mixed effect models 

Host AFDW.
```{r}
#### P acuta -- this needs to be done with Pacuta_stress_data b/c recovery timepoints are small or not existent.
## Host 
Pacuta_Host_LMER <- lmer(Host_AFDW.mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Pacuta_Host_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Pacuta_Host_LMER)
qqPlot(residuals(Pacuta_Host_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_Host_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$Host_AFDW.mg.cm2 ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_Host_LMER, ddf="lme4", type='III')
Anova(Pacuta_Host_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Host_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

Sym AFDW. 
```{r}
## Sym
Pacuta_Sym_LMER <- lmer(Sym_AFDW.mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Pacuta_Sym_LMER)
qqPlot(residuals(Pacuta_Sym_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_Sym_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$Sym_AFDW.mg.cm2 ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_Sym_LMER, ddf="lme4", type='III')
Anova(Pacuta_Sym_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Sym_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

S:H Ratio AFDW 
```{r}
## S:H Ratio 
Pacuta_SHRatio_LMER <- lmer(log(Ratio_AFDW.mg.cm2) ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Pacuta_SHRatio_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Pacuta_SHRatio_LMER)
qqPlot(residuals(Pacuta_SHRatio_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_SHRatio_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$Ratio_AFDW.mg.cm2 ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_SHRatio_LMER, ddf="lme4", type='III')
Anova(Pacuta_SHRatio_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_SHRatio_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

*M. capitata* mixed effect models. 

Host AFDW. 
```{r}
## Host 
Mcap_Host_LMER <- lmer(Host_AFDW.mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Mcap_Host_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Mcap_Host_LMER)
qqPlot(residuals(Mcap_Host_LMER)) # qqplot and histogram 
hist(residuals(Mcap_Host_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$Host_AFDW.mg.cm2 ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_Host_LMER, ddf="lme4", type='III')
Anova(Mcap_Host_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Host_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

Sym AFDW. 
```{r}
## Sym
Mcap_Sym_LMER <- lmer(Sym_AFDW.mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Mcap_Sym_LMER)
qqPlot(residuals(Mcap_Sym_LMER)) # qqplot and histogram 
hist(residuals(Mcap_Sym_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$Sym_AFDW.mg.cm2 ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_Sym_LMER, ddf="lme4", type='III')
Anova(Mcap_Sym_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Sym_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

S:H Ratio AFDW. 
```{r}
## S:H Ratio 
Mcap_SHRatio_LMER <- lmer(Ratio_AFDW.mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Mcap_SHRatio_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Mcap_SHRatio_LMER)
qqPlot(residuals(Mcap_SHRatio_LMER)) # qqplot and histogram 
hist(residuals(Mcap_SHRatio_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$Ratio_AFDW.mg.cm2 ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_SHRatio_LMER, ddf="lme4", type='III')
Anova(Mcap_SHRatio_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_SHRatio_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

### Soluble Protein

Raw distribution view. 
```{r}
hist(pacuta_full_data$prot_mg.cm2) 
hist(mcap_full_data$prot_mg.cm2) 
# both slightly right skewed positively 
```

Mixed effect modeling. 

*P. acuta*
```{r}
## Pacuta 
Pacuta_prot_LMER <- lmer(prot_mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

## summary of the model chosen from above
## plot(Mcap_Chl_LMER) in console to view diagnostic plots
summary(Pacuta_prot_LMER)
qqPlot(residuals(Pacuta_prot_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_prot_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$prot_mg.cm2 ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2)

## significance of model
anova(Pacuta_prot_LMER, ddf="lme4", type='III')
Anova(Pacuta_prot_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_prot_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

*M. capitata* 
```{r}
## Mcap
Mcap_prot_LMER <- lmer(prot_mg.cm2 ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Mcap_prot_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_prot_LMER) in console to view diagnostic plots
summary(Mcap_prot_LMER)
qqPlot(residuals(Mcap_prot_LMER)) # qqplot and histogram 
hist(residuals(Mcap_prot_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$prot_mg.cm2 ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2)

## significance of model
anova(Mcap_prot_LMER, ddf="lme4", type='III')
Anova(Mcap_prot_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_prot_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

### Host Total Antioxidant Capacity

Raw distribution view. 
```{r}
hist(pacuta_full_data$cre.umol.mgprot) 
hist(mcap_full_data$cre.umol.mgprot) 
# both very right skewed positively 
```

Mixed effect modeling  

*P. acuta*
```{r}
## Pacuta 
Pacuta_TAC_LMER <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Pacuta_TAC_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_prot_LMER) in console to view diagnostic plots
summary(Pacuta_TAC_LMER)
qqPlot(residuals(Pacuta_TAC_LMER)) # qqplot and histogram 
hist(residuals(Pacuta_TAC_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta_full_data$cre.umol.mgprot ~ pacuta_full_data$Timepoint*pacuta_full_data$Temperature*pacuta_full_data$CO2) 

## significance of model
anova(Pacuta_TAC_LMER, ddf="lme4", type='III')
Anova(Pacuta_TAC_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_TAC_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

*M. capitata*
```{r}
## Mcapitata
Mcap_TAC_LMER <- lmer(log(cre.umol.mgprot) ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=mcap_full_data) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

isSingular(Mcap_TAC_LMER, tol = 1e-4) # this produced TRUE, the GLMM is overfitting the model

## summary of the model chosen from above
## plot(Mcap_TAC_LMER) in console to view diagnostic plots
summary(Mcap_TAC_LMER)
qqPlot(residuals(Mcap_TAC_LMER)) # qqplot and histogram 
hist(residuals(Mcap_TAC_LMER)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap_full_data$cre.umol.mgprot ~ mcap_full_data$Timepoint*mcap_full_data$Temperature*mcap_full_data$CO2) 

## significance of model
anova(Mcap_TAC_LMER, ddf="lme4", type='III')
Anova(Mcap_TAC_LMER, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_TAC_LMER, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

### Bleaching Score 

Raw distribution score.
```{r}
color.data$Plug_ID <- as.character(color.data$Plug_ID)
color.data$Blch.Time <- as.factor(color.data$Blch.Time)
mcap.color <- color.data %>% subset(Species == "Mcapitata")
pacuta.color <- color.data %>% subset(Species == "Pacuta")

hist(pacuta.color$Bleaching.Score) 
hist(mcap.color$Bleaching.Score) 
# both very left/negatively skewed 
```


Mixed effect modeling. Adding Plug ID as random factor because this was repeated measures. 
```{r}
## Mcap 
mcap.color$Color.Transformed <- mcap.color$Bleaching.Score+100
range(mcap.color$Color.Transformed)
hist(mcap.color$Color.Transformed)

Mcap_Color_LMER_tank_plug <- lmer((Bleaching.Score^(1/4)) ~ Blch.Time*Temperature*CO2 + (1|Tank/Plug_ID), na.action=na.omit, data=mcap.color) #GLMM; nesting Plug ID in tank because the Plug IDs never moved tanks 
# GAM 
# anova(GLMM, GAM) # report AIC values
Mcap_Color_LMER_tank_plug <- lmer(log(Bleaching.Score) ~ Blch.Time*Temperature*CO2 + (1|Tank/Plug_ID), na.action=na.omit, data=mcap.color)

## summary of the model chosen from above
## plot(Mcap_TAC_LMER) in console to view diagnostic plots
summary(Mcap_Color_LMER_tank_plug)
qqPlot(residuals(Mcap_Color_LMER_tank_plug)) # qqplot and histogram 
hist(residuals(Mcap_Color_LMER_tank_plug)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap.color$Bleaching.Score ~ mcap.color$Timepoint*mcap.color$Temperature*mcap.color$CO2)

## significance of model
anova(Mcap_Color_LMER_tank_plug, ddf="lme4", type='III')
Anova(Mcap_Color_LMER_tank_plug, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Color_LMER_tank_plug, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

```{r}
## Pacuta 
pacuta.color$Color.Transformed <- pacuta.color$Bleaching.Score+120
range(pacuta.color$Color.Transformed)
hist(pacuta.color$Color.Transformed)

Pacuta_Color_LMER_tank_plug <- lmer(log(Color.Transformed) ~ Blch.Time*Temperature*CO2 + (1|Tank/Plug_ID), na.action=na.omit, data=pacuta.color) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values
 

alias(Pacuta_Color_GLM) # to see which groups are issues and being dropped as columns / coefficients 

## summary of the model chosen from above
## plot(Mcap_TAC_LMER) in console to view diagnostic plots
summary(Pacuta_Color_LMER_tank_plug)
qqPlot(residuals(Pacuta_Color_LMER_tank_plug)) # qqplot and histogram 
hist(residuals(Pacuta_Color_LMER_tank_plug)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta.color$Color.Transformed ~ pacuta.color$Timepoint*pacuta.color$Temperature*pacuta.color$CO2)

## significance of model
anova(Pacuta_Color_LMER_tank_plug, ddf="lme4", type='III')
Anova(Pacuta_Color_LMER_tank_plug, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Color_LMER_tank_plug, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

### Buoyant Weight 

Raw distribution score.
```{r}
Growth.data$Plug_ID <- as.character(Growth.data$Plug_ID)
Growth.data$Growth.Time <- as.factor(Growth.data$Growth.Time)
mcap.Growth <- Growth.data %>% subset(Species == "Mcapitata")
pacuta.Growth <- Growth.data %>% subset(Species == "Pacuta")

hist(pacuta.Growth$Growth.Rate) 
hist(mcap.Growth$Growth.Rate) 
# both very heavy tailed distribution
```

*M. capitata* mixed effect modeling.
```{r}
#  excluded in BW script (6 and 8 week 1210 Mcapitata - ATAC Tank 1)

Mcap_Growth_LMER_tank_plug <- lmer(Growth.Rate ~ Growth.Time*Temperature*CO2 + (1|Tank/Plug_ID), na.action=na.omit, data=mcap.Growth) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

## summary of the model chosen from above
## plot(Mcap_TAC_LMER) in console to view diagnostic plots
summary(Mcap_Growth_LMER_tank_plug)
qqPlot(residuals(Mcap_Growth_LMER_tank_plug)) # qqplot and histogram 
hist(residuals(Mcap_Growth_LMER_tank_plug)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(mcap.Growth$Growth.Rate ~ mcap.Growth$Timepoint*mcap.Growth$Temperature*mcap.Growth$CO2)

## significance of model
anova(Mcap_Growth_LMER_tank_plug, ddf="lme4", type='III')
Anova(Mcap_Growth_LMER_tank_plug, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Mcap_Growth_LMER_tank_plug, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```

*P. acuta* mixed effect modeling.
```{r}
Pacuta_Growth_LMER_tank_plug <- lmer(log(Growth.Rate) ~ Growth.Time*Temperature*CO2 + (1|Tank/Plug_ID), na.action=na.omit, data=pacuta.Growth) #GLMM
# GAM 
# anova(GLMM, GAM) # report AIC values

## summary of the model chosen from above
## plot(Mcap_TAC_LMER) in console to view diagnostic plots
summary(Pacuta_Growth_LMER_tank_plug)
qqPlot(residuals(Pacuta_Growth_LMER_tank_plug)) # qqplot and histogram 
hist(residuals(Pacuta_Growth_LMER_tank_plug)) 
## checking homogeneity of variance using levene test (can also use Bartlett's test function)
leveneTest(pacuta.Growth$Growth.Rate ~ pacuta.Growth$Timepoint*pacuta.Growth$Temperature*pacuta.Growth$CO2)

## significance of model
anova(Pacuta_Growth_LMER_tank_plug, ddf="lme4", type='III')
Anova(Pacuta_Growth_LMER_tank_plug, ddf="lme4", type='III') # Type III Wald chisquare tests

anova(Pacuta_Growth_LMER_tank_plug, ddf="Satterthwaite", type='III') # Type III Analysis of Variance Table with Satterthwaite's method
```


# Multivariate analysis 

## PCA dataframes and color choices 
```{r}
## Stress PCAs 
pca_data <- data[complete.cases(data), ] # only taking rows where all samples have data
Pacuta_data <- subset(pca_data, Species=="Pacuta") # subset to P. acuta
Mcap_data <- subset(pca_data, Species=="Mcapitata") # subset to M. capitata

Mcap_scaled_data <- prcomp(Mcap_data[c(10:19)], scale=TRUE, center=TRUE) # columns 10-19 are the data values
Pacuta_scaled_data <- prcomp(Pacuta_data[c(10:19)], scale=TRUE, center=TRUE) # columns 10-19 are the data values

## Screeplot
# P acuta 
Pacuta_data.cov <- cov(Pacuta_data[,c(10:19)])
Pacuta_data.eigen <- eigen(Pacuta_data.cov)
plot(Pacuta_data.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Pacuta_data.eigen$values)
# Mcap  
Mcap_data.cov <- cov(Mcap_data[,c(10:19)])
Mcap_data.eigen <- eigen(Mcap_data.cov)
plot(Mcap_data.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Mcap_data.eigen$values)

## Timeseries PCAs 
pca_timeseriesdata <- timeseries.data[complete.cases(timeseries.data), ] # only taking rows where all samples have data
Pacuta_timeseriesdata <- subset(pca_timeseriesdata, Species=="Pacuta") %>% subset(Group != "Recovery HTAC") # subset to P. acuta
Mcap_timeseriesdata <- subset(pca_timeseriesdata, Species=="Mcapitata") # subset to M. capitata
  Mcap_acute_data <- subset(Mcap_timeseriesdata, Phase=="Acute")

Mcap_scaled_timeseriesdata <- prcomp(Mcap_timeseriesdata[c(10:16)], scale=TRUE, center=TRUE) # columns 8-14 are the data values
Pacuta_scaled_timeseriesdata <- prcomp(Pacuta_timeseriesdata[c(10:16)], scale=TRUE, center=TRUE) # columns 10-16 are the data values

## Screeplot
## P acuta 
Pacuta_timeseriesdata.cov <- cov(Pacuta_timeseriesdata[,c(10:16)])
Pacuta_timeseriesdata.eigen <- eigen(Pacuta_timeseriesdata.cov)
plot(Pacuta_timeseriesdata.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Pacuta_timeseriesdata.eigen$values)
# M cap
Mcap_timeseriesdata.cov <- cov(Mcap_timeseriesdata[,c(10:16)])
Mcap_timeseriesdata.eigen <- eigen(Mcap_timeseriesdata.cov)
plot(Mcap_timeseriesdata.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Mcap_timeseriesdata.eigen$values)
```

## Plotting stress timepoints. 
```{r}
# Choosing colors for PCA 
line.cols <- c("lightblue", "blue", "salmon", "red3", "lightblue", "blue", "salmon", "red3")
fill.cols <- c("white", "white", "white", "white", "lightblue", "blue", "salmon", "red3")

Mcap_group <- autoplot(Mcap_scaled_data, data = Mcap_data, colour = 'Group', shape = "Phase",
         loadings = TRUE, loadings.colour = 'black', size=1, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1,
         frame=TRUE, frame.type = 'norm') + theme_bw() +
  scale_color_manual(values = line.cols) + scale_fill_manual(values = fill.cols) + ggtitle("M. capitata"); Mcap_group

Pacuta_group <- autoplot(Pacuta_scaled_data, data = Pacuta_data, colour = 'Group', shape = "Phase", frame=TRUE, frame.type = 'norm',
         loadings = TRUE, loadings.colour = 'black', size=1, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + 
  theme(legend.position = "top") +
  theme_bw() + scale_color_manual(values = line.cols) + scale_fill_manual(values = fill.cols) + ggtitle("P. acuta"); Pacuta_group

Stress.PCAs <- arrangeGrob(Mcap_group, Pacuta_group, ncol=1)
ggsave(file="Output/Final_Figures/PCA-stress.pdf", Stress.PCAs, width = 8, height = 10, units = c("in"))
ggsave(file="Output/Final_Figures/PCA-stress.png", Stress.PCAs, width = 8, height = 10, units = c("in"))
```

## Plotting timeseries with recovery timepoints. 
```{r}
# Choosing colors for PCA 
time.line.cols <- c("lightblue", "blue", "salmon", "red3", "lightblue", "blue", "salmon", "red3", "lightblue", "blue", "salmon", "red3")
time.fill.cols <- c("white", "white", "white", "white", "lightblue", "blue", "salmon", "red3", "dodgerblue2", "navyblue", "firebrick3", "darkred")
plines.cols <- c("lightblue", "blue", "salmon", "red3", "lightblue", "blue", "salmon", "red3", "lightblue", "blue")
pfill.cols <- c("white", "white", "white", "white", "lightblue", "blue", "salmon", "red3", "dodgerblue2", "navyblue")
  
Pacuta_timeseries <- autoplot(Pacuta_scaled_timeseriesdata, data = Pacuta_timeseriesdata, colour = 'Group', shape = "Phase",
         loadings = TRUE, loadings.colour = 'black', size=1, loadings.label = TRUE, frame=TRUE, frame.type = 'norm',
         loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + 
  theme_bw() + scale_color_manual(values = plines.cols) + scale_fill_manual(values = pfill.cols) + ggtitle("P. acuta"); Pacuta_timeseries

Mcap_timeseries <- autoplot(Mcap_scaled_timeseriesdata, data = Mcap_timeseriesdata, colour = 'Group', shape = "Phase",
         loadings = TRUE, loadings.colour = 'black', size=1, loadings.label = TRUE, 
         loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1,
         frame=TRUE, frame.type = 'norm') + theme_bw() +
  scale_color_manual(values = time.line.cols) + scale_fill_manual(values = time.fill.cols) + ggtitle("M. capitata"); Mcap_timeseries

Timeseries.PCAs <- arrangeGrob(Mcap_timeseries, Pacuta_timeseries, ncol=1)
ggsave(file="Output/Final_Figures/PCA-timeseries.pdf", Timeseries.PCAs, width = 8, height = 10, units = c("in"))
ggsave(file="Output/Final_Figures/PCA-timeseries.png", Timeseries.PCAs, width = 8, height = 10, units = c("in"))
```

### 3 phases 
Start with pca_data subsetted to each species 

```{r}
mcap.three <- autoplot(Mcap_scaled_timeseriesdata, data = Mcap_timeseriesdata, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'norm',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  theme(legend.position = "none") +
  facet_grid(~ Phase) + theme_bw() +
  scale_color_manual(values = cols) + scale_fill_manual(values = cols) + ggtitle("Montipora capitata"); mcap.three

pacuta.three <- autoplot(Pacuta_scaled_timeseriesdata, data = Pacuta_timeseriesdata, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'norm',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  theme(legend.position = "none") +
  facet_grid(~ Phase) + theme_bw() +
  scale_color_manual(values = cols) + scale_fill_manual(values = cols) + ggtitle("Pocillopora acuta"); pacuta.three

Three <- arrangeGrob(mcap.three, pacuta.three, ncol=1)
ggsave(file="Output/Final_Figures/PCA-threephase.pdf", Three, width = 12, height = 12, units = c("in"))
ggsave(file="Output/Final_Figures/PCA-threephase.png", Three, width = 12, height = 12, units = c("in"))
```

## PERMANOVA Analysis 

For P. acuta and M. cap. timeseries (without resp and photo). 
```{r}
## Pacuta - effect of treatment and timepoint
# scale data
pacuta_timeseries_vegan <- scale(Pacuta_timeseriesdata[ ,10:16])
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(pacuta_timeseries_vegan ~ Timepoint*Temperature*CO2, data = Pacuta_timeseriesdata, method='eu')
adonis(pacuta_timeseries_vegan ~ Phase*Temperature*CO2, data = Pacuta_timeseriesdata, method='eu')

## Mcap - effect of treatment and timepoint
# scale data
mcap_timeseries_vegan <- scale(Mcap_timeseriesdata[ ,10:16])
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(mcap_timeseries_vegan ~ Timepoint*Temperature*CO2, data = Mcap_timeseriesdata, method='eu')
adonis(mcap_timeseries_vegan ~ Phase*Temperature*CO2, data = Mcap_timeseriesdata, method='eu')
```

