---
title: "Statistics.Rmd"
author: "EL Strand"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clears workspace. 
```{r}
rm(list=ls())
```

Loading libraries. 
```{r}
library(ggplot2)
library(tidyverse)
library(plyr)
library(dplyr)
library(car) # qqplot and ANOVA function
library(mgcv) # GAM function
library(ggfortify)
library(devtools)
install_github("vqv/ggbiplot")
library(lme4)
library(gridExtra)
library(vegan)
library(effects)
```

### Read in data files to analyze. 

```{r}
Master <- read.csv("Environmental_data/Master_Fragment.csv") %>% select(Plug_ID, Species, ANALYSIS, Timepoint, Treatment, Tank, Temperature, CO2) %>%
  filter(ANALYSIS == "Physiology" | ANALYSIS == "Physiology/Molecular")
  
Photo.Resp <- read.csv("Physiology_variables/Respirometry/Output/Resp_AllRates.csv") %>% select(Fragment.ID.x, Pnet_umol.cm2.hr, Rdark_umol.cm2.hr, Pgross_umol.cm2.hr) %>%
  dplyr::rename(Plug_ID = Fragment.ID.x)
  Photo.Resp$Plug_ID <- gsub("P-", "", Photo.Resp$Plug_ID)
  Photo.Resp$Plug_ID <- gsub("M-", "", Photo.Resp$Plug_ID)
  Photo.Resp <- Photo.Resp %>% separate(Plug_ID, c("Plug_ID", "Date")) %>% select(-Date)

Chlorophyll <-  read.csv("Output/chlorophyll.csv") %>% select(Plug_ID, chla.ug.cm2, chlc2.ug.cm2)

Biomass <- read.csv("Output/Tissue_Biomass.csv") %>% select(Plug_ID, partner, AFDW.mg.cm2)

Protein <- read.csv("Output/soluble_protein.csv") %>% select(Plug_ID, prot_mg.cm2)

Antioxidant <- read.csv("Output/antioxidant_capacity.csv") %>% select(Plug_ID, cre.umol.mgprot)

Color.Score <- read.csv("Physiology_variables/Bleaching_Score_Output.csv") %>% select(PLUG.ID, Bleaching.Score, Timepoint) %>% dplyr::rename(Plug_ID = PLUG.ID) %>% 
  dplyr::rename(Blch.Time = Timepoint)
Color.Score$Blch.Time <- gsub("^.{0,4}", "", Color.Score$Blch.Time) # taking out the "Week" in every row
Color.Score$Blch.Time <- paste(Color.Score$Blch.Time, "week") # adding "week" after each number in every row 

Growth <- read.csv("Physiology_variables/Growth_Buoyant_Weight/Growth_Rates.csv") %>% select(PLUG.ID, T0.g.d.cm2, T1.g.d.cm2, T2.g.d.cm2, T3.g.d.cm2, T4.g.d.cm2, T5.g.d.cm2,
                                                                                             T6.g.d.cm2, T7.g.d.cm2, T8.g.d.cm2) %>% 
  gather("Growth.Time", "Growth.Rate", 2:10) %>% na.omit(Growth.Rate) %>% dplyr::rename(Plug_ID=PLUG.ID) %>%
  mutate(Growth.Time = case_when(
    Growth.Time == "T0.g.d.cm2" ~ "1 week",
    Growth.Time == "T1.g.d.cm2" ~ "2 week", 
    Growth.Time == "T2.g.d.cm2" ~ "4 week",
    Growth.Time == "T3.g.d.cm2" ~ "6 week",
    Growth.Time == "T4.g.d.cm2" ~ "8 week",
    Growth.Time == "T5.g.d.cm2" ~ "10 week",
    Growth.Time == "T6.g.d.cm2" ~ "12 week",
    Growth.Time == "T7.g.d.cm2" ~ "14 week",
    Growth.Time == "T8.g.d.cm2" ~ "16 week")
  )


# Cell.density

# come back to adding Color.Score and Growth back in and subsetting for where all three timepoint columns match 
# joining all datasets together, reshaping the AFDW partner into columns, renaming those columns, removing an NA column, and calculating S:H AFDW ratio
data <- join_all(list(Master, Photo.Resp, Chlorophyll, Biomass, Protein, Antioxidant), by='Plug_ID', type='left') %>%
  spread(partner, AFDW.mg.cm2) %>% dplyr::rename(Host_AFDW.mg.cm2 = Host) %>% dplyr::rename(Sym_AFDW.mg.cm2 = Sym) %>% select(-"<NA>") %>%
  mutate(Ratio_AFDW.mg.cm2=Sym_AFDW.mg.cm2/Host_AFDW.mg.cm2) 
data$Timepoint <- factor(data$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week", "12 week", "16 week"))

timeseries.data <- data %>% select(-Pnet_umol.cm2.hr, -Rdark_umol.cm2.hr, -Pgross_umol.cm2.hr)

pacuta_full_data <- data %>% subset(Species == "Pacuta")
mcap_full_data <- data %>% subset(Species == "Mcapitata")
```

# Univariate analysis

## Statistical Modeling
Fixed factors: Temperature, CO2, Species, Timepoint. 
Random factors: Tank
Come back to this: Should I be doing this separately for species? I think so.. 

### Photosynthetic Rates and Respiration 

Raw data distribution view.
```{r}
hist(pacuta_full_data$Pgross_umol.cm2.hr)
hist(pacuta_full_data$Pnet_umol.cm2.hr)
hist(pacuta_full_data$Rdark_umol.cm2.hr)

hist(mcap_full_data$Pgross_umol.cm2.hr)
hist(mcap_full_data$Pnet_umol.cm2.hr)
hist(mcap_full_data$Rdark_umol.cm2.hr)
# mcap doesn't look normal, the models should account for this but might need to transform post models 
```

P.acuta linear mixed effect models
```{r}
## Pnet
Pacuta_Pnet_LMER_tank <- lmer(Pnet_umol.cm2.hr ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
Pacuta_Pnet_GLM <- glm(Pnet_umol.cm2.hr ~ Timepoint*Temperature*CO2, na.action=na.omit, data=pacuta_full_data) # GLM
anova(Pacuta_Pnet_LMER_tank, Pacuta_Pnet_GLM) # AIC with tank = 151.40; without tank = 146.81. GLMM (5 points lower) with tank wins, move on to residuals

summary(Pacuta_Pnet_LMER_tank)
qqPlot(residuals(Pacuta_Pnet_LMER_tank))
hist(residuals(Pacuta_Pnet_LMER_tank)) 
# the above looks normal; no need to transform 

Anova(Pacuta_Pnet_LMER_tank, type='III') #Anova needs to be capital A to be from car function and distinguish types of ANOVAs 
plot(allEffects(Pacuta_Pnet_LMER_tank))

## Pgross 
Pacuta_Pgross_LMER_tank <- lmer(Pgross_umol.cm2.hr ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
Pacuta_Pgross_GLM <- glm(Pgross_umol.cm2.hr ~ Timepoint*Temperature*CO2, na.action=na.omit, data=pacuta_full_data) # GLM
anova(Pacuta_Pgross_LMER_tank, Pacuta_Pgross_GLM) # AIC with tank = 196.04; without tank = 199.33. GLMM (3 points lower) with tank wins, move on to residuals. 

summary(Pacuta_Pgross_LMER_tank)
qqPlot(residuals(Pacuta_Pgross_LMER_tank))
hist(residuals(Pacuta_Pgross_LMER_tank)) 
# the above looks normal; no need to transform 

Anova(Pacuta_Pgross_LMER_tank, type='III') #Anova needs to be capital A to be from car function and distinguish types of ANOVAs 
plot(allEffects(Pacuta_Pgross_LMER_tank))

## Dark Resp
Pacuta_Resp_LMER_tank <- lmer(Rdark_umol.cm2.hr ~ Timepoint*Temperature*CO2 + (1|Tank), na.action=na.omit, data=pacuta_full_data) #GLMM
Pacuta_Resp_GLM <- glm(Rdark_umol.cm2.hr ~ Timepoint*Temperature*CO2, na.action=na.omit, data=pacuta_full_data) # GLM
anova(Pacuta_Resp_LMER_tank, Pacuta_Resp_GLM) # AIC with tank = -293.73; without tank = -295.44. The difference is <2 points, so there is no real difference between models so the less complex model will be better = the GLM 

summary(Pacuta_Resp_GLM)
qqPlot(residuals(Pacuta_Resp_GLM))
hist(residuals(Pacuta_Resp_GLM)) 
# This looks a bit skewed, but how skewed is too skewed? Come back to this 
```

Generalized Additive model 
```{r}
Pnet_GAM <- gam(Pnet_umol.cm2.hr ~ s(Treatment) + s(Species) + s(Temperature) + s(CO2), data=data)
summary(Pnet_GAM)
## getting error with smoothing function (s(Timepoint)) b/c might be having trouble with distance of factors? come back to this
## Error in names(dat) <- object$term : 'names' attribute [1] must be the same length as the vector [0]

anova(Pnet_LMER_tank, Pnet_GAM)
```


# Multivariate analysis 

## PCA 

Plot PCA with prcomp to center and scale. 

```{r}
pca_data <- data[complete.cases(data), ] # only taking rows where all samples have data

scaled_data <- prcomp(pca_data[c(9:18)], scale=TRUE, center=TRUE) # columns 8-17 are the data values

## screeplot
pca_data.cov <- cov(pca_data[,c(9:18)])
pca_data.eigen <- eigen(pca_data.cov)
plot(pca_data.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(pca_data.eigen$values)
```

Group by species, timepoint, treatment. 
```{r}
cols <- c("lightblue", "blue", "salmon", "red3")

Species.PCA <- autoplot(scaled_data, data = pca_data, colour = 'Species',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.5); Species.PCA

Treatment.PCA <- autoplot(scaled_data, data = pca_data, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.5) + scale_color_manual(values = cols) + scale_fill_manual(values = cols); Treatment.PCA

Timepoint.PCA <- autoplot(scaled_data, data = pca_data, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.5) + scale_colour_brewer(); Timepoint.PCA

ggsave(file="Output/Final_Figures/General-Species-PCA.pdf", Species.PCA, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/General-Species-PCA.png", Species.PCA, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/General-Treatment-PCA.pdf", Treatment.PCA, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/General-Treatment-PCA.png", Treatment.PCA, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/General-Timepoint-PCA.pdf", Timepoint.PCA, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/General-Timepoint-PCA.png", Timepoint.PCA, width = 11, height = 6, units = c("in"))

General.PCAs <- arrangeGrob(Species.PCA, Treatment.PCA, Timepoint.PCA, ncol=2)
ggsave(file="Output/Final_Figures/General-PCAs.pdf", General.PCAs, width = 10, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/General-PCAs.png", General.PCAs, width = 10, height = 6, units = c("in"))
```

Subsetted to only P. acuta
```{r}
Pacuta_data <- subset(pca_data, Species=="Pacuta")
Pacuta_scaled_data <- prcomp(Pacuta_data[c(9:18)], scale=TRUE, center=TRUE) # columns 8-17 are the data values

## screeplot
Pacuta_data.cov <- cov(Pacuta_data[,c(9:18)])
Pacuta_data.eigen <- eigen(Pacuta_data.cov)
plot(Pacuta_data.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Pacuta_data.eigen$values)

Pacuta_treatment <- autoplot(Pacuta_scaled_data, data = Pacuta_data, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.25) + scale_color_manual(values = cols) + scale_fill_manual(values = cols) + ggtitle("P. acuta"); Pacuta_treatment

Pacuta_timepoint <- autoplot(Pacuta_scaled_data, data = Pacuta_data, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.25) + scale_colour_brewer() + ggtitle("P. acuta"); Pacuta_timepoint

ggsave(file="Output/Final_Figures/Pacuta-Treatment-PCA.pdf", Pacuta_treatment, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Pacuta-Treatment-PCA.png", Pacuta_treatment, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/Pacuta-Timepoint-PCA.pdf", Pacuta_timepoint, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Pacuta-Timepoint-PCA.png", Pacuta_timepoint, width = 11, height = 6, units = c("in"))

Pacuta.stress.PCAs <- arrangeGrob(Pacuta_treatment, Pacuta_timepoint, ncol=2)
ggsave(file="Output/Final_Figures/Pacuta-stress-PCAs.pdf", Pacuta.stress.PCAs, width = 10, height = 4, units = c("in"))
ggsave(file="Output/Final_Figures/Pacuta-stress-PCAs.png", Pacuta.stress.PCAs, width = 10, height = 4, units = c("in"))
```

Subsetted to M. capitata
```{r}
Mcap_data <- subset(pca_data, Species=="Mcapitata")
Mcap_scaled_data <- prcomp(Mcap_data[c(9:18)], scale=TRUE, center=TRUE) # columns 8-17 are the data values

## screeplot
Mcap_data.cov <- cov(Mcap_data[,c(9:18)])
Mcap_data.eigen <- eigen(Mcap_data.cov)
plot(Mcap_data.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Mcap_data.eigen$values)

Mcap_treatment <- autoplot(Mcap_scaled_data, data = Mcap_data, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim() + 
  scale_color_manual(values = cols) + scale_fill_manual(values = cols) + ggtitle("M. capitata"); Mcap_treatment

Mcap_timepoint <- autoplot(Mcap_scaled_data, data = Mcap_data, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  
  #xlim(-0.4,0.4) + 
  scale_colour_brewer() + ggtitle("M. capitata"); Mcap_timepoint

ggsave(file="Output/Final_Figures/Mcap-Treatment-PCA.pdf", Mcap_treatment, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Mcap-Treatment-PCA.png", Mcap_treatment, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/Mcap-Timepoint-PCA.pdf", Mcap_timepoint, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Mcap-Timepoint-PCA.png", Mcap_timepoint, width = 11, height = 6, units = c("in"))

Mcap.stress.PCAs <- arrangeGrob(Mcap_treatment, Mcap_timepoint, ncol=2)
ggsave(file="Output/Final_Figures/Mcap-stress-PCAs.pdf", Mcap.stress.PCAs, width = 10, height = 4, units = c("in"))
ggsave(file="Output/Final_Figures/Mcap-stress-PCAs.png", Mcap.stress.PCAs, width = 10, height = 4, units = c("in"))
```
PERMANOVA for P. acuta and M. cap. 
Stress treatments (with resp and photo)
```{r}
## Pacuta - effect of treatment and timepoint
# scale data
pacuta_vegan <- scale(Pacuta_data[ ,9:18])
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(pacuta_vegan ~ Treatment*Timepoint, data = Pacuta_data, method='eu')

## Mcap - effect of treatment and timepoint
# scale data
mcap_vegan <- scale(Mcap_data[ ,9:18])
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(mcap_vegan ~ Treatment*Timepoint, data = Mcap_data, method='eu')
```


Entire timeseries (without resp/photo and with recovery timepoints).
```{r}
pca_timeseriesdata <- timeseries.data[complete.cases(timeseries.data), ] # only taking rows where all samples have data
timeseries_scaled_data <- prcomp(pca_timeseriesdata[c(9:15)], scale=TRUE, center=TRUE) # columns 8-14 are the data values

## screeplot
pca_timeseriesdata.cov <- cov(pca_timeseriesdata[,c(9:15)])
pca_timeseriesdata.eigen <- eigen(pca_timeseriesdata.cov)
plot(pca_timeseriesdata.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(pca_timeseriesdata.eigen$values)

Timeseries.Species.PCA <- autoplot(timeseries_scaled_data, data = pca_timeseriesdata, colour = 'Species',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw(); Timeseries.Species.PCA

Timeseries.Treatment.PCA <- autoplot(timeseries_scaled_data, data = pca_timeseriesdata, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim(-0.23,0.13) + 
  scale_color_manual(values = cols) + scale_fill_manual(values = cols); Timeseries.Treatment.PCA

Timeseries.Timepoint.PCA <- autoplot(timeseries_scaled_data, data = pca_timeseriesdata, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim(-0.23,0.13) + 
  scale_colour_brewer(); Timeseries.Timepoint.PCA

ggsave(file="Output/Final_Figures/Timeseries-Species-PCA.pdf", Timeseries.Species.PCA, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Timeseries-Species-PCA.png", Timeseries.Species.PCA, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/Timeseries-Treatment-PCA.pdf", Timeseries.Treatment.PCA, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Timeseries-Treatment-PCA.png", Timeseries.Treatment.PCA, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/Timeseries-Timepoint-PCA.pdf", Timeseries.Timepoint.PCA, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Timeseries-Timepoint-PCA.png", Timeseries.Timepoint.PCA, width = 11, height = 6, units = c("in"))

General.Timeseries.PCAs <- arrangeGrob(Timeseries.Species.PCA, Timeseries.Treatment.PCA, Timeseries.Timepoint.PCA, ncol=2)
ggsave(file="Output/Final_Figures/General-timeseries-PCAs.pdf", General.Timeseries.PCAs, width = 10, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/General-timeseries-PCAs.png", General.Timeseries.PCAs, width = 10, height = 6, units = c("in"))
```

Entire timeseries subsetted to only P. acuta
```{r}
Pacuta_timeseriesdata <- subset(pca_timeseriesdata, Species=="Pacuta")
Pacuta_scaled_timeseriesdata <- prcomp(Pacuta_timeseriesdata[c(9:15)], scale=TRUE, center=TRUE) # columns 8-14 are the data values

## screeplot
Pacuta_timeseriesdata.cov <- cov(Pacuta_timeseriesdata[,c(9:15)])
Pacuta_timeseriesdata.eigen <- eigen(Pacuta_timeseriesdata.cov)
plot(Pacuta_timeseriesdata.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Pacuta_timeseriesdata.eigen$values)

Pacuta_timeseries_treatment <- autoplot(Pacuta_scaled_timeseriesdata, data = Pacuta_timeseriesdata, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim(-0.225,0.2) + 
  scale_color_manual(values = cols) + scale_fill_manual(values = cols) + ggtitle("P. acuta"); Pacuta_timeseries_treatment

Pacuta_timeseries_timepoint <- autoplot(Pacuta_scaled_timeseriesdata, data = Pacuta_timeseriesdata, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim(-0.225,0.2) + 
  scale_colour_brewer() + ggtitle("P. acuta"); Pacuta_timeseries_timepoint

ggsave(file="Output/Final_Figures/Pacuta-Timeseries-Treatment-PCA.pdf", Pacuta_timeseries_treatment, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Pacuta-Timeseries-Treatment-PCA.png", Pacuta_timeseries_treatment, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/Pacuta-Timeseries-Timepoint-PCA.pdf", Pacuta_timeseries_timepoint, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Pacuta-Timeseries-Timepoint-PCA.png", Pacuta_timeseries_timepoint, width = 11, height = 6, units = c("in"))

Pacuta.timeseries.PCAs <- arrangeGrob(Pacuta_timeseries_treatment, Pacuta_timeseries_timepoint, ncol=2)
ggsave(file="Output/Final_Figures/Pacuta-Timeseries-PCAs.pdf", Pacuta.timeseries.PCAs, width = 10, height = 4, units = c("in"))
ggsave(file="Output/Final_Figures/Pacuta-Timeseries-PCAs.png", Pacuta.timeseries.PCAs, width = 10, height = 4, units = c("in"))
```
Entire timeseries only subsetted to M. capitata
```{r}
Mcap_timeseriesdata <- subset(pca_timeseriesdata, Species=="Mcapitata")
Mcap_scaled_timeseriesdata <- prcomp(Mcap_timeseriesdata[c(9:15)], scale=TRUE, center=TRUE) # columns 8-14 are the data values

## screeplot
Mcap_timeseriesdata.cov <- cov(Mcap_timeseriesdata[,c(9:15)])
Mcap_timeseriesdata.eigen <- eigen(Mcap_timeseriesdata.cov)
plot(Mcap_timeseriesdata.eigen$values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph') + lines(Mcap_timeseriesdata.eigen$values)

Mcap_timeseries_treatment <- autoplot(Mcap_scaled_timeseriesdata, data = Mcap_timeseriesdata, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim(-0.225,0.13) + 
  scale_color_manual(values = cols) + scale_fill_manual(values = cols) + ggtitle("M. capitata"); Mcap_timeseries_treatment

Mcap_timeseries_timepoint <- autoplot(Mcap_scaled_timeseriesdata, data = Mcap_timeseriesdata, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  #xlim(-0.225,0.13) + 
  scale_colour_brewer() + ggtitle("M. capitata"); Mcap_timeseries_timepoint

ggsave(file="Output/Final_Figures/Mcap-Timeseries-Treatment-PCA.pdf", Mcap_timeseries_treatment, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Mcap-Timeseries-Treatment-PCA.png", Mcap_timeseries_treatment, width = 11, height = 6, units = c("in"))

ggsave(file="Output/Final_Figures/Mcap-Timeseries-Timepoint-PCA.pdf", Mcap_timeseries_timepoint, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Mcap-Timeseries-Timepoint-PCA.png", Mcap_timeseries_timepoint, width = 11, height = 6, units = c("in"))

Mcap.timeseries.PCAs <- arrangeGrob(Mcap_timeseries_treatment, Mcap_timeseries_timepoint, ncol=2)
ggsave(file="Output/Final_Figures/Mcap-Timeseries-PCAs.pdf", Mcap.timeseries.PCAs, width = 10, height = 4, units = c("in"))
ggsave(file="Output/Final_Figures/Mcap-Timeseries-PCAs.png", Mcap.timeseries.PCAs, width = 10, height = 4, units = c("in"))
```

PERMANOVA for P. acuta and M. cap. 
Timeseries (without resp and photo)
```{r}
## Pacuta - effect of treatment and timepoint
# scale data
pacuta_timeseries_vegan <- scale(Pacuta_timeseriesdata[ ,9:15])
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(pacuta_timeseries_vegan ~ Treatment*Timepoint, data = Pacuta_timeseriesdata, method='eu')

## Mcap - effect of treatment and timepoint
# scale data
mcap_timeseries_vegan <- scale(Mcap_timeseriesdata[ ,9:15])
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis(mcap_timeseries_vegan ~ Treatment*Timepoint, data = Mcap_timeseriesdata, method='eu')
```
