---
title: "Statistics.Rmd"
author: "EL Strand"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clears workspace. 
```{r}
rm(list=ls())
```

Loading libraries. 
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(car) # qqplot function
library(mgcv) # GAM function
library(ggfortify)
library(devtools)
install_github("vqv/ggbiplot")
```

### Read in data files to analyze. 

```{r}
Master <- read.csv("Environmental_data/Master_Fragment.csv") %>% select(Plug_ID, Species, ANALYSIS, Timepoint, Treatment, Temperature, CO2) %>%
  filter(ANALYSIS == "Physiology" | ANALYSIS == "Physiology/Molecular")
  
Photo.Resp <- read.csv("Physiology_variables/Respirometry/Output/Resp_AllRates.csv") %>% select(Fragment.ID.x, Pnet_umol.cm2.hr, Rdark_umol.cm2.hr, Pgross_umol.cm2.hr) %>%
  dplyr::rename(Plug_ID = Fragment.ID.x)
  Photo.Resp$Plug_ID <- gsub("P-", "", Photo.Resp$Plug_ID)
  Photo.Resp$Plug_ID <- gsub("M-", "", Photo.Resp$Plug_ID)
  Photo.Resp <- Photo.Resp %>% separate(Plug_ID, c("Plug_ID", "Date")) %>% select(-Date)

Chlorophyll <-  read.csv("Output/chlorophyll.csv") %>% select(Plug_ID, chla.ug.cm2, chlc2.ug.cm2)

Biomass <- read.csv("Output/Tissue_Biomass.csv") %>% select(Plug_ID, partner, AFDW.mg.cm2)

Protein <- read.csv("Output/soluble_protein.csv") %>% select(Plug_ID, prot_mg.cm2)

Antioxidant <- read.csv("Output/antioxidant_capacity.csv") %>% select(Plug_ID, cre.umol.mgprot)

Color.Score <- read.csv("Physiology_variables/Bleaching_Score_Output.csv") %>% select(PLUG.ID, Bleaching.Score, Timepoint) %>% rename(Plug_ID=PLUG.ID) %>% 
  rename(Blch.Time = Timepoint)
Color.Score$Blch.Time <- gsub("^.{0,4}", "", Color.Score$Blch.Time) # taking out the "Week" in every row
Color.Score$Blch.Time <- paste(Color.Score$Blch.Time, "week") # adding "week" after each number in every row 

Growth <- read.csv("Physiology_variables/Growth_Buoyant_Weight/Growth_Rates.csv") %>% select(PLUG.ID, T0.g.d.cm2, T1.g.d.cm2, T2.g.d.cm2, T3.g.d.cm2, T4.g.d.cm2, T5.g.d.cm2,
                                                                                             T6.g.d.cm2, T7.g.d.cm2, T8.g.d.cm2) %>% 
  gather("Growth.Time", "Growth.Rate", 2:10) %>% na.omit(Growth.Rate) %>% rename(Plug_ID=PLUG.ID) %>%
  mutate(Growth.Time = case_when(
    Growth.Time == "T0.g.d.cm2" ~ "1 week",
    Growth.Time == "T1.g.d.cm2" ~ "2 week", 
    Growth.Time == "T2.g.d.cm2" ~ "4 week",
    Growth.Time == "T3.g.d.cm2" ~ "6 week",
    Growth.Time == "T4.g.d.cm2" ~ "8 week",
    Growth.Time == "T5.g.d.cm2" ~ "10 week",
    Growth.Time == "T6.g.d.cm2" ~ "12 week",
    Growth.Time == "T7.g.d.cm2" ~ "14 week",
    Growth.Time == "T8.g.d.cm2" ~ "16 week")
  )


# Cell.density

# come back to adding Color.Score and Growth back in and subsetting for where all three timepoint columns match 
# joining all datasets together, reshaping the AFDW partner into columns, renaming those columns, removing an NA column, and calculating S:H AFDW ratio
data <- join_all(list(Master, Photo.Resp, Chlorophyll, Biomass, Protein, Antioxidant), by='Plug_ID', type='left') %>%
  spread(partner, AFDW.mg.cm2) %>% rename(Host_AFDW.mg.cm2 = Host) %>% rename(Sym_AFDW.mg.cm2 = Sym) %>% select(-"<NA>") %>%
  mutate(Ratio_AFDW.mg.cm2=Sym_AFDW.mg.cm2/Host_AFDW.mg.cm2) 
data$Timepoint <- factor(data$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week"))

```

# Univariate analysis

## Statistical Modeling
Fixed factors: Temperature, CO2, Species, Timepoint. 
Come back to this: Should I be doing this separately for species? I think so.. 

### Photosynthetic Rates and Respiration 

Linear mixed effect model
```{r}
Pnet_LMER <- lmer(Pnet_umol.cm2.hr ~ Timepoint*Species*Temperature*CO2 + (1|Plug_ID), na.action=na.omit, data=data)
summary(Pnet_LMER)

qqPlot(residuals(Pnet_LMER))
hist(residuals(Pnet_LMER))
```

Generalized Additive model 
```{r}
Pnet_GAM <- gam(Pnet_umol.cm2.hr ~ s(Timepoint) + s(Species) + s(Temperature) + s(CO2), data=data)
summary(Pnet_GAM)


```

Comparing these two models. 
```{r}
anova(mod_lm2, mod_gam2, test="Chisq")
```



# Multivariate analysis 

## PCA 

Plot PCA with prcomp to center and scale. 

```{r}
pca_data <- data[complete.cases(data), ] # only taking rows where all samples have data

scaled_data <- prcomp(pca_data[c(8:17)], scale=TRUE, center=TRUE) # columns 8-17 are the data values
```

Group by species, timepoint, treatment. 
```{r}
cols <- c("lightblue", "blue", "salmon", "red3")

Species.PCA <- autoplot(scaled_data, data = pca_data, colour = 'Species',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.5); Species.PCA

Treatment.PCA <- autoplot(scaled_data, data = pca_data, colour = 'Treatment',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.5) + scale_color_manual(values = cols) + scale_fill_manual(values = cols); Treatment.PCA

Timepoint.PCA <- autoplot(scaled_data, data = pca_data, colour = 'Timepoint',
         loadings = TRUE, loadings.colour = 'black', frame=TRUE, size=0.5, frame.type = 'convex',
         loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = 'black', loadings.label.vjust=-1) + theme_bw() +
  xlim(-0.25,0.5) + scale_colour_brewer(); Timepoint.PCA
```



