---
title: "Tissue_Biomass"
author: "EL Strand"
date: "7/18/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")
if (!require("dplyr")) install.packages("dplyr")
if (!require("plyr")) install.packages("plyr")
if (!require("Rmisc")) install.packages("Rmisc")

# load packages
library(tidyverse)
library(plotrix)
library(dplyr)
library(plyr)
library(Rmisc)
```

Dry Weight - organic and inorganic compounds
Ash-Free Dry Weight - organic compounds 

Read in Ash-Free Dry Weight raw data, surface area, homogenate volume, and master datasheets
```{r}
AFDW_raw <- read.csv("Physiology_variables/Tissue_Biomass/AFDW_data.csv") %>% filter(!is.na(Plug_ID))

# Load surface area and homogenate volume values; merge dataframes from chla, chlc means 
Surface_area <- read.csv("Physiology_variables/Respirometry/coral_surface_area.csv") %>% dplyr::rename(Plug_ID = Sample_ID) %>%
  select(Plug_ID, surface.area.cm2)
Homogenate_volume <- read.csv("Physiology_variables/homogenate_vol.csv") %>%
  select(Plug_ID, vol_mL)
Master_fragment <- read.csv("Environmental_data/Master_Fragment_20200616_QC.csv") %>%
  select(Plug_ID, Species, Treatment, Temperature, CO2, Sample.Date)
  Master_fragment$Plug_ID <- as.character(Master_fragment$Plug_ID)
  Master_fragment$Sample.Date <- as.character(Master_fragment$Sample.Date)
Sampling_days <- read.csv("Environmental_data/Sampling_dates.csv") %>% dplyr::rename(Sample.Date = Date)
  Sampling_days$Sample.Date <- as.character(Sampling_days$Sample.Date)

AFDW_raw <- join_all(list(AFDW_raw, Surface_area, Homogenate_volume, Master_fragment), by='Plug_ID', type='left')
AFDW_raw <- full_join(AFDW_raw, Sampling_days, by="Sample.Date", type='left')
```

Calculate mass per mL
```{r}
#different volumes for sym (4ml) and host (5ml)
sym <- 5
host <- 4
AFDW <- AFDW_raw %>%
  mutate(dry.pan.mass.g.ml = case_when(partner=="Host" ~ dry.pan.mass.g/sym, partner=="Sym" ~dry.pan.mass.g/host),
         burnt.pan.mass.g.ml = case_when(partner=="Host" ~ burnt.pan.mass.g/sym, partner=="Sym" ~burnt.pan.mass.g/host))
```

Calculating Tissue Biomass (AFDW)
```{r}
# Calculating Dry Weight and Ash-Free Dry Weight; standardizing by homogenate volume and surface area
AFDW <- AFDW %>%
  mutate(dry.bioimass.g.ml = (dry.pan.mass.g.ml - burnt.pan.mass.g.ml),
         dry.bioimass.g = dry.pan.mass.g.ml * vol_mL,
         burnt.bioimass.g = burnt.pan.mass.g.ml * vol_mL,
         DW.mg.cm2 = ((dry.bioimass.g)*1000) / surface.area.cm2,
         AFDW.mg.cm2 = ((dry.bioimass.g-burnt.bioimass.g)*1000) / surface.area.cm2)
AFDW

# Transforming the dataframe to be easily analyzed and visualized by type of measurement
AFDW_final <- AFDW %>% select(Plug_ID, partner, Species, Treatment, Temperature, CO2, Sample.Date, Day_Number, DW.mg.cm2, AFDW.mg.cm2) %>%
  gather("Measurement", "Value", 9:10) %>%
  filter(!is.na(Value))

## temporarily taking out samples that will need to be re-done
AFDW_final <- AFDW_final %>% filter(Value > 0)
AFDW_final

# Calculating means per treatment and timepoint. 
AFDW_means_dayno <- summarySE(AFDW_final, measurevar="Value", groupvars=c("partner", "Day_Number", "Treatment", "Temperature", "CO2", "Species", "Measurement")) %>%
  subset(Measurement=="AFDW.mg.cm2")
AFDW_means_dayno

AFDW_means_date <- summarySE(AFDW_final, measurevar="Value", groupvars=c("partner", "Sample.Date", "Treatment", "Temperature", "CO2", "Species", "Measurement")) %>%
  subset(Measurement=="AFDW.mg.cm2")
AFDW_means_date
```

Graphing data.
```{r}
cols <- c("lightblue", "blue", "salmon", "red3")

ggplot(AFDW_final, aes(x=Treatment, y=Value, fill=partner)) + geom_boxplot(alpha=0.5) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + xlab("") + theme_bw() + 
  theme(legend.position = "left") + 
  facet_grid(Measurement ~ Species, scales = "free") 
```

```{r}
ggplot(AFDW_means_dayno, aes(x=Day_Number, y = Value, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(partner ~ Species, scales = "free") +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Number of Days") + 
  ylab("Ash-Free Dry Weight (mg.cm2)") + # y-axis label and y axis range
  theme_bw() + 
  ggtitle("A) Ash-Free Dry Weight") +
  theme(legend.position = "none") +
  geom_vline(xintercept = c(58), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)) #Set the text angle

ggplot(AFDW_means_date, aes(x=Sample.Date, y = Value, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(partner ~ Species, scales = "free") +
  geom_errorbar(aes(ymin=Value-se, ymax=Value+se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Sample Date") + 
  ylab("Ash-Free Dry Weight (mg.cm2)") + # y-axis label and y axis range
  theme_bw() + 
  ggtitle("A) Ash-Free Dry Weight") +
  theme(legend.position = "none") +
  geom_vline(xintercept = c(58), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5)) #Set the text angle
```

############## Bleaching Fragment Test prior to running the rest of the samples. 

Reading in bleaching test information. 
```{r}
Phys_testing <- read.csv("Physiology_variables/phys_testing.csv")
Phys_testing$colony_id <- as.character(Phys_testing$colony_id)
AFDW$colony_id <- as.character(AFDW$colony_id)

AFDW_testing <- full_join(AFDW, Phys_testing, by="colony_id") %>% 
select(colony_id, partner, Species, TREATMENT, Sample.Date, DW.mg.cm2, AFDW.mg.cm2, status, bleaching_score) %>% filter(!is.na(status))

AFDW_testing <- gather(AFDW_testing, "Measurement", "Value", 6:7) # changing columns to rows to allow for easier facet_grid in graphing section
```

Graphing bleach test data. 
```{r}
Bleached1 <- ggplot(AFDW_testing, aes(x=status, y=Value, fill=partner)) + geom_boxplot(alpha=0.5) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + xlab("") + theme_bw() + 
  theme(legend.position = "left") + 
  facet_grid(Measurement ~ Species, scales = "free")
Bleached1

Bleached2 <- ggplot(AFDW_testing, aes(x=bleaching_score, y=Value, color=partner)) + geom_point() + 
  xlab("Bleaching Score") + theme_bw() + 
  facet_grid(Measurement ~ Species, scales = "free") + 
  theme(legend.position = "left")
Bleached2

Bleached.Figs <- combine_plots(Bleached1, Bleached2,
  labels = c("A", "B"), ncol=2)

ggsave(file="Output/Phystest_AFDW.pdf", Bleached.Figs, width = 11, height = 10, units = c("in"))
```


