---
title: "Total_antioxidant_capacity"
author: "EL Strand"
date: "8/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clears workspace. 
```{r}
rm(list=ls())
```

Load in required libraries. 
```{r}
# load packages
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(broom)
library(Rmisc)
```

Based on Total Antioxidant Capacity kit found in protocols folder. 

Reading in datafiles. 
```{r}
## 20201014 
# Plate 1 
Oct14plate1_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201014_plate1_TAC_platemap.csv") 
  Oct14plate1_platemap$Plug_ID <- as.character(Oct14plate1_platemap$Plug_ID)
Oct14_plate1_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201014_plate1_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Oct14_plate1_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201014_plate1_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Oct14_plate1 <- full_join(Oct14plate1_platemap, Oct14_plate1_initial, by="Well") %>%
  full_join(Oct14_plate1_final) %>%
  filter(!is.na(Plug_ID))

# Plate 3 
Oct14plate3_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201014_plate3_TAC_platemap.csv") 
  Oct14plate3_platemap$Plug_ID <- as.character(Oct14plate3_platemap$Plug_ID)
Oct14_plate3_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201014_plate3_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Oct14_plate3_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201014_plate3_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Oct14_plate3 <- full_join(Oct14plate3_platemap, Oct14_plate3_initial, by="Well") %>%
  full_join(Oct14_plate3_final) %>%
  filter(!is.na(Plug_ID))

Oct14 <- union(Oct14_plate1, Oct14_plate3) # %>% union(Oct14_plate3)
Oct14_st <- Oct14 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID)

## 20201025 Plates 7, 8, 9
# Plate 1 
Oct25plate1_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate1_TAC_platemap.csv") 
  Oct25plate1_platemap$Plug_ID <- as.character(Oct25plate1_platemap$Plug_ID)
Oct25_plate1_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate1_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Oct25_plate1_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate1_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Oct25_plate1 <- full_join(Oct25plate1_platemap, Oct25_plate1_initial, by="Well") %>%
  full_join(Oct25_plate1_final) %>%
  filter(!is.na(Plug_ID))

# Plate 2
Oct25plate2_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate2_TAC_platemap.csv") 
  Oct25plate2_platemap$Plug_ID <- as.character(Oct25plate2_platemap$Plug_ID)
Oct25_plate2_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate2_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Oct25_plate2_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate2_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Oct25_plate2 <- full_join(Oct25plate2_platemap, Oct25_plate2_initial, by="Well") %>%
  full_join(Oct25_plate2_final) %>%
  filter(!is.na(Plug_ID))

# Plate 3
Oct25plate3_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate3_TAC_platemap.csv") 
  Oct25plate3_platemap$Plug_ID <- as.character(Oct25plate3_platemap$Plug_ID)
Oct25_plate3_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate3_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Oct25_plate3_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201025_plate3_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Oct25_plate3 <- full_join(Oct25plate3_platemap, Oct25_plate3_initial, by="Well") %>%
  full_join(Oct25_plate3_final) %>%
  filter(!is.na(Plug_ID))

Oct25 <- union(Oct25_plate1, Oct25_plate2) %>% union(Oct25_plate3)
Oct25_st <- Oct25 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID)

## 20201110 
Nov10plate1_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201110_plate1_TAC_platemap.csv") 
  Nov10plate1_platemap$Plug_ID <- as.character(Nov10plate1_platemap$Plug_ID)
Nov10_plate1_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201110_plate1_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Nov10_plate1_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201110_plate1_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Nov10 <- full_join(Nov10plate1_platemap, Nov10_plate1_initial, by="Well") %>%
  full_join(Nov10_plate1_final) %>%
  filter(!is.na(Plug_ID))

Nov10_st <- Nov10 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID)

## 20201118 
Nov18plate1_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201118_plate1_TAC_platemap.csv") 
  Nov18plate1_platemap$Plug_ID <- as.character(Nov18plate1_platemap$Plug_ID)
Nov18_plate1_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201118_plate1_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Nov18_plate1_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201118_plate1_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Nov18_plate1 <- full_join(Nov18plate1_platemap, Nov18_plate1_initial, by="Well") %>%
  full_join(Nov18_plate1_final) %>%
  filter(!is.na(Plug_ID))

Nov18plate2_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201118_plate2_TAC_platemap.csv") 
  Nov18plate2_platemap$Plug_ID <- as.character(Nov18plate2_platemap$Plug_ID)
Nov18_plate2_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201118_plate2_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Nov18_plate2_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20201118_plate2_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Nov18_plate2 <- full_join(Nov18plate2_platemap, Nov18_plate2_initial, by="Well") %>%
  full_join(Nov18_plate2_final) %>%
  filter(!is.na(Plug_ID))

Nov18 <- union(Nov18_plate1, Nov18_plate2)
Nov18_st <- Nov18 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID)

## 20210215 
# Plate 1
Feb15plate1_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate1_TAC_platemap.csv") 
  Feb15plate1_platemap$Plug_ID <- as.character(Feb15plate1_platemap$Plug_ID)
Feb15_plate1_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate1_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Feb15_plate1_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate1_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Feb15_plate1 <- full_join(Feb15plate1_platemap, Feb15_plate1_initial, by="Well") %>%
  full_join(Feb15_plate1_final) %>%
  filter(!is.na(Plug_ID))

# Plate 1
Feb15plate2_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate2_TAC_platemap.csv") 
  Feb15plate2_platemap$Plug_ID <- as.character(Feb15plate2_platemap$Plug_ID)
Feb15_plate2_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate2_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Feb15_plate2_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate2_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Feb15_plate2 <- full_join(Feb15plate2_platemap, Feb15_plate2_initial, by="Well") %>%
  full_join(Feb15_plate2_final) %>%
  filter(!is.na(Plug_ID))

# Plate 3
Feb15plate3_platemap <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate3_TAC_platemap.csv") 
  Feb15plate3_platemap$Plug_ID <- as.character(Feb15plate3_platemap$Plug_ID)
Feb15_plate3_final <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate3_final.csv", header=T) %>% dplyr::rename(Abs_final = X490.490)
Feb15_plate3_initial <- read.csv("Physiology_variables/Total_Antioxidant_Capacity/20210215_plate3_initial.csv", header=T) %>% dplyr::rename(Abs_initial = X490.490)
Feb15_plate3 <- full_join(Feb15plate3_platemap, Feb15_plate3_initial, by="Well") %>%
  full_join(Feb15_plate3_final) %>%
  filter(!is.na(Plug_ID))

Feb15 <- union(Feb15_plate1, Feb15_plate2) %>% union(Feb15_plate3)
Feb15_st <- Feb15 %>% filter(grepl("Standard", Plug_ID)) %>% dplyr::rename(Standard = Plug_ID)
```

Checking for outliers, duplicates, and effect of plate. 
```{r}
QC <- union(Oct14, Oct25) %>% union(Nov10) %>% union(Nov18) %>% union(Feb15)
boxplot(QC$Abs_final)
hist(QC$Abs_final)

Duplicates <- QC %>% dplyr::group_by(Plug_ID) %>% 
  dplyr::summarise(n = n()) %>% filter(n > 2)
Duplicates

filter(QC, Plug_ID == "Standard10")
filter(Feb15, Plug_ID == "1046")
## output is list of samples that were run more than once; each standard should have the equal number of days run (Oct7, 14, 25, Nov18) x2 replicates
# Not reading B7 Standard 10 on 20201118 plate on QC for some reason
```

Plot standard curve. 
```{r}
st.name <- c("Standard1", "Standard2", "Standard3", "Standard4", "Standard5", "Standard6", "Standard7", "Standard8", "Standard9", "Standard10")
concentration <- c("1", "0.5", "0.25", "0.125", "0.0625", "0.03125", "0.01560", "0.00780", "0.00390", "0")

## 20201014
Oct14_st <- Oct14_st %>%
  mutate(Abs_net = Abs_final - Abs_initial) %>% 
  dplyr::group_by(Standard) %>%
  summarise(Abs_net = mean(Abs_net)) %>%
  mutate(Abs_net.adj = Abs_net - Abs_net[Standard == "Standard10"])

Oct14_st$conc_mM <- concentration[match(Oct14_st$Standard, st.name)]
  Oct14_st$conc_mM <- as.numeric(Oct14_st$conc_mM)

TAC_mod_lin3 <- lm(conc_mM ~ Abs_net.adj, data = Oct14_st)
summary(TAC_mod_lin3)

std_curve_plot3 <- Oct14_st %>%
  ggplot(aes(x = Abs_net.adj, y = conc_mM)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  xlim(0,1.5) +
  labs(title = paste("Adj R2 = ", signif(summary(TAC_mod_lin3)$adj.r.squared, 5),
                     "; TAC Standard Curve Oct 14 Standards")) + 
  geom_smooth(method = "lm")
std_curve_plot3

## 20201025
Oct25_st <- Oct25_st %>%
  mutate(Abs_net = Abs_final - Abs_initial) %>% 
  dplyr::group_by(Standard) %>%
  summarise(Abs_net = mean(Abs_net)) %>%
  mutate(Abs_net.adj = Abs_net - Abs_net[Standard == "Standard10"])

Oct25_st$conc_mM <- concentration[match(Oct25_st$Standard, st.name)]
  Oct25_st$conc_mM <- as.numeric(Oct25_st$conc_mM)

TAC_mod_lin4 <- lm(conc_mM ~ Abs_net.adj, data = Oct25_st)
summary(TAC_mod_lin4)

std_curve_plot4 <- Oct25_st %>%
  ggplot(aes(x = Abs_net.adj, y = conc_mM)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  xlim(0,1.5) +
  labs(title = paste("Adj R2 = ", signif(summary(TAC_mod_lin4)$adj.r.squared, 5),
                     "; TAC Standard Curve Oct 25 Standards")) + 
  geom_smooth(method = "lm")
std_curve_plot4

## 20201110
Nov10_st <- Nov10_st %>%
  mutate(Abs_net = Abs_final - Abs_initial) %>% 
  dplyr::group_by(Standard) %>%
  summarise(Abs_net = mean(Abs_net)) %>%
  mutate(Abs_net.adj = Abs_net - Abs_net[Standard == "Standard10"])

Nov10_st$conc_mM <- concentration[match(Nov10_st$Standard, st.name)]
  Nov10_st$conc_mM <- as.numeric(Nov10_st$conc_mM)

TAC_mod_linNov10 <- lm(conc_mM ~ Abs_net.adj, data = Nov10_st)
summary(TAC_mod_linNov10)

std_curve_plotNov10 <- Nov10_st %>%
  ggplot(aes(x = Abs_net.adj, y = conc_mM)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  xlim(0,1.5) +
  labs(title = paste("Adj R2 = ", signif(summary(TAC_mod_linNov10)$adj.r.squared, 5),
                     "; TAC Standard Curve Nov 10 Standards")) + 
  geom_smooth(method = "lm")
std_curve_plotNov10

## 20201118
Nov18_st <- Nov18_st %>%
  mutate(Abs_net = Abs_final - Abs_initial) %>% 
  dplyr::group_by(Standard) %>%
  summarise(Abs_net = mean(Abs_net)) %>%
  mutate(Abs_net.adj = Abs_net - Abs_net[Standard == "Standard10"])

Nov18_st$conc_mM <- concentration[match(Nov18_st$Standard, st.name)]
  Nov18_st$conc_mM <- as.numeric(Nov18_st$conc_mM)

TAC_mod_linNov18 <- lm(conc_mM ~ Abs_net.adj, data = Nov18_st)
summary(TAC_mod_linNov18)

std_curve_plotNov18 <- Nov18_st %>%
  ggplot(aes(x = Abs_net.adj, y = conc_mM)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  xlim(0,1.5) +
  labs(title = paste("Adj R2 = ", signif(summary(TAC_mod_linNov18)$adj.r.squared, 5),
                     "; TAC Standard Curve Nov 18 Standards")) + 
  geom_smooth(method = "lm")
std_curve_plotNov18

## 20210215
Feb15_st <- Feb15_st %>%
  mutate(Abs_net = Abs_final - Abs_initial) %>% 
  dplyr::group_by(Standard) %>%
  summarise(Abs_net = mean(Abs_net)) %>%
  mutate(Abs_net.adj = Abs_net - Abs_net[Standard == "Standard10"])

Feb15_st$conc_mM <- concentration[match(Feb15_st$Standard, st.name)]
  Feb15_st$conc_mM <- as.numeric(Feb15_st$conc_mM)

TAC_mod_linFeb15 <- lm(conc_mM ~ Abs_net.adj, data = Feb15_st)
summary(TAC_mod_linFeb15)

std_curve_plotFeb15 <- Feb15_st %>%
  ggplot(aes(x = Abs_net.adj, y = conc_mM)) +
  geom_point(color = "blue", size = 1.5) + 
  theme_bw() + 
  xlim(0,1.5) +
  labs(title = paste("Adj R2 = ", signif(summary(TAC_mod_linFeb15)$adj.r.squared, 5),
                     "; TAC Standard Curve Feb 15 Standards")) + 
  geom_smooth(method = "lm")
std_curve_plotFeb15
```


```{r}
## All points projected on Nov 10 standard curve plot (std_curve_plotNov18) ; standard curve 5 (Nov10_st) ; linear model 5 (TAC_mod_linNov10)

## October 14 Plates (3)
Oct14 <- Oct14 %>% 
  mutate(Abs_net = Abs_final - Abs_initial) %>% #subtract initial from final to get delta abs
  dplyr::group_by(Plug_ID) %>% # group to calculate mean
  summarise(Abs_net = mean(Abs_net)) %>% #calculate mean abs of duplicates
  mutate(Abs_net.adj = Abs_net - Nov10_st$Abs_net[Nov10_st$Standard == "Standard10"],
         uae.mM = map_dbl(Abs_net.adj, ~ predict(TAC_mod_linNov10, newdata = data.frame(Abs_net.adj = .)))) #use the regression to calculate all the mean UAE values from the abs values
std_curve_plotNov10 + #plot the standard curve to check this against the protocol
  geom_point(data = Oct14, aes(x = Abs_net.adj, y = uae.mM), pch = "X", cex = 3, alpha = 0.3) +
    labs(title = "Oct 14 samples projected on Nov 10 curve")

## October 25 Plates (3)
Oct25 <- Oct25 %>% 
  mutate(Abs_net = Abs_final - Abs_initial) %>% #subtract initial from final to get delta abs
  dplyr::group_by(Plug_ID) %>% # group to calculate mean
  summarise(Abs_net = mean(Abs_net)) %>% #calculate mean abs of duplicates
  mutate(Abs_net.adj = Abs_net - Nov10_st$Abs_net[Nov10_st$Standard == "Standard10"],
         uae.mM = map_dbl(Abs_net.adj, ~ predict(TAC_mod_linNov10, newdata = data.frame(Abs_net.adj = .)))) #use the regression to calculate all the mean UAE values from the abs values
std_curve_plotNov10 + #plot the standard curve to check this against the protocol
  geom_point(data = Oct25, aes(x = Abs_net.adj, y = uae.mM), pch = "X", cex = 3, alpha = 0.3) +
    labs(title = "Oct 25 samples projected on Nov 10 curve")

## November 10 Plate
Nov10 <- Nov10 %>% 
  mutate(Abs_net = Abs_final - Abs_initial) %>% #subtract initial from final to get delta abs
  dplyr::group_by(Plug_ID) %>% # group to calculate mean
  summarise(Abs_net = mean(Abs_net)) %>% #calculate mean abs of duplicates
  mutate(Abs_net.adj = Abs_net - Nov10_st$Abs_net[Nov10_st$Standard == "Standard10"],
         uae.mM = map_dbl(Abs_net.adj, ~ predict(TAC_mod_linNov10, newdata = data.frame(Abs_net.adj = .)))) #use the regression to calculate all the mean UAE values from the abs values
std_curve_plotNov10 + #plot the standard curve to check this against the protocol
  geom_point(data = Nov10, aes(x = Abs_net.adj, y = uae.mM), pch = "X", cex = 3, alpha = 0.3) +
    labs(title = "Nov10 samples projected on Nov 10 curve")

## November 18 Plates
Nov18 <- Nov18 %>% 
  mutate(Abs_net = Abs_final - Abs_initial) %>% #subtract initial from final to get delta abs
  dplyr::group_by(Plug_ID) %>% # group to calculate mean
  summarise(Abs_net = mean(Abs_net)) %>% #calculate mean abs of duplicates
  mutate(Abs_net.adj = Abs_net - Nov10_st$Abs_net[Nov10_st$Standard == "Standard10"],
         uae.mM = map_dbl(Abs_net.adj, ~ predict(TAC_mod_linNov10, newdata = data.frame(Abs_net.adj = .)))) #use the regression to calculate all the mean UAE values from the abs values
std_curve_plotNov10 + #plot the standard curve to check this against the protocol
  geom_point(data = Nov18, aes(x = Abs_net.adj, y = uae.mM), pch = "X", cex = 3, alpha = 0.3) +
    labs(title = "Nov18 samples projected on Nov 10 curve")

## February 15 Plates
Feb15 <- Feb15 %>% 
  mutate(Abs_net = Abs_final - Abs_initial) %>% #subtract initial from final to get delta abs
  dplyr::group_by(Plug_ID) %>% # group to calculate mean
  summarise(Abs_net = mean(Abs_net)) %>% #calculate mean abs of duplicates
  mutate(Abs_net.adj = Abs_net - Nov10_st$Abs_net[Nov10_st$Standard == "Standard10"],
         uae.mM = map_dbl(Abs_net.adj, ~ predict(TAC_mod_linNov10, newdata = data.frame(Abs_net.adj = .)))) #use the regression to calculate all the mean UAE values from the abs values
std_curve_plotNov10 + #plot the standard curve to check this against the protocol
  geom_point(data = Nov18, aes(x = Abs_net.adj, y = uae.mM), pch = "X", cex = 3, alpha = 0.3) +
    labs(title = "Feb15 samples projected on Nov 10 curve")
```

Samples above highest standard point for that plate need to be diluted and re-done.  
```{r}
## October 14 plates
filter(Oct14, Plug_ID == "Standard1") # uae.mM is 1.004828
Oct14dilute.tac <- Oct14 %>% filter(uae.mM > 1.004828) # output is 0 samples

## October 25 plates
filter(Oct25, Plug_ID == "Standard1") # uae.mM is 1.005419	
Oct25dilut.tac <- Oct25 %>% filter(uae.mM > 1.005419) # output is 1075
## 1075 re-done on Nov10 plate

## November 10 plate 
filter(Nov10, Plug_ID == "Standard1") # uae.mM is 0.995335	
Nov10dilut.tac <- Nov10 %>% filter(uae.mM > 0.995335)   # output is 0 samples

## November 18 plate 
filter(Nov18, Plug_ID == "Standard1") # uae.mM is 0.6954394	
Nov18dilut.tac <- Nov18 %>% filter(uae.mM > 0.6954395)   # output is 0 samples

## February 15 plate 
filter(Feb15, Plug_ID == "Standard1") # uae.mM is 0.7837479		
Feb15dilut.tac <- Feb15 %>% filter(uae.mM > 0.7837479)   # output is 0 samples

# combine list 
Dilute.TAC.list <- union(Oct14dilute.tac, Oct25dilut.tac) %>% union(Nov10dilut.tac) %>% union(Nov18dilut.tac) %>%
  union(Feb15dilut.tac)
Dilute.TAC.list

##output as of 20210118 = none to re-do. 
```

Reading in to surface area, soluble protein, and homogenate volume. 
```{r}
sa <- read_csv("Physiology_variables/Respirometry/coral_surface_area.csv") %>% dplyr::rename(Plug_ID = Sample_ID) %>% select(Plug_ID, surface.area.cm2)
  sa$Plug_ID <- as.character(sa$Plug_ID)
Homogenate_volume <- read.csv("Physiology_variables/homogenate_vol.csv") %>% select(Plug_ID, vol_mL)
  Homogenate_volume$Plug_ID <- as.character(Homogenate_volume$Plug_ID)

Master_fragment <- read.csv("Environmental_data/Master_Fragment.csv", header=T, sep=",", na.string="NA") %>% select(Plug_ID, Species, Timepoint, Temperature, Treatment, CO2)
  Master_fragment$Plug_ID <- as.character(Master_fragment$Plug_ID)

prot <- read_csv("output/soluble_protein.csv") 
  prot$Plug_ID <- as.character(prot$Plug_ID)

metadata <- full_join(sa, Homogenate_volume, by='Plug_ID', type='left') %>% 
  full_join(Master_fragment) %>% full_join(prot) 
```

Calculate TAC. 
```{r}
# combine Oct14, Oct25 files, Nov10, Nov18, Feb15; filter out standard values
TAC.values <- union(Oct14, Oct25) %>% union(Nov10) %>% union(Nov18) %>% union(Feb15)
TAC.values <- TAC.values[!grepl("Standard", TAC.values$Plug_ID),] 

TAC <- left_join(TAC.values, metadata) %>%
   mutate(cre.umol.L = uae.mM * 2189) %>%   # Convert to CRE (see product manual) per unit sample volume
   mutate(cre.umol = cre.umol.L * (vol_mL / 1000),  # Convert to CRE per coral by multiplying by homog. vol.
          cre.umol.mgprot = cre.umol / (prot_ug / 1000))  # Convert to CRE per mg protein by dividing by total protein

range(TAC$cre.umol.mgprot)
```

Writing TAC output file. Taking out P1488 (See notes in results package markdown).
```{r}
# taking out sample 1488 
TAC <- TAC %>% filter(!Plug_ID == "1488")
filter(TAC, Plug_ID == "1488") # checking to see that worked; output should be 0

TAC %>% select(Plug_ID, uae.mM, Species, Timepoint, Treatment, Temperature, CO2, prot_ug, cre.umol.L, cre.umol, cre.umol.mgprot) %>%
  write_csv(., file = "Output/antioxidant_capacity.csv")
```

Calculating mean values per treatment per time point. Checking sample sizes. Reading file back in to avoid having to run the above code multiple times. 
```{r}
TAC <- read.csv("Output/antioxidant_capacity.csv") %>% na.omit()

TAC.values <- ggplot(TAC, aes(x=Treatment, y=cre.umol.mgprot, color = Species)) + geom_boxplot() + theme_classic() 
ggsave(file="Output/Final_Figures/TAC-values.png", TAC.values, width = 11, height = 10, units = c("in"))

TAC.means <- summarySE(TAC, measurevar="cre.umol.mgprot", groupvars=c("Timepoint", "Treatment", "Temperature", "CO2", "Species")) %>% filter(N>2) %>%
  mutate(Species = case_when(
    Species == "Mcapitata" ~ "Montipora capitata",
    Species == "Pacuta" ~ "Pocillopora acuta"
  ))
TAC.means$Timepoint <- factor(TAC.means$Timepoint, levels=c("Day 1","Day 2","1 week","2 week","4 week","6 week","8 week","12 week","16 week"))

# Looking at sample sizes 
dplyr::filter(TAC.means, N > 6) # this output should be zero b/c the biggest sample size is 6
dplyr::filter(TAC.means, N < 6) # this output should just be the Day 1, Day 2, Week 1, 12 week, and 16 week decreased sample sizes
```

Plotting over time and treatment. 
```{r}
cols <- c("lightblue", "blue", "pink", "red")

TAC.plot <- ggplot(TAC.means, aes(x=Timepoint, y = cre.umol.mgprot, group = Treatment, color = Treatment)) +
  geom_line() + geom_point() +
  facet_grid(~Species) +
  geom_errorbar(aes(ymin=cre.umol.mgprot-se, ymax=cre.umol.mgprot+se), width=.2) +
  scale_color_manual(values = cols) + 
  xlab("Timepoint") + 
  theme_classic() + 
  ylab(expression("Copper Reducing Equivalents" ~(umol ~mg^-1))) +
  #ggtitle("Total Antioxidant Capacity") +
  theme(legend.position = "left") +
  geom_vline(xintercept = c(7.2), colour="black", linetype="dotted") +
  theme(plot.title = element_text(face = 'bold.italic', size = 12, hjust = 0)) + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.2, hjust = 1.3)) + #Set the text angle
  theme(panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change axis line
  axis.line = element_line(colour = "black")); TAC.plot

ggsave(file="Output/Final_Figures/Total-Antioxidant-Capacity.pdf", TAC.plot, width = 11, height = 6, units = c("in"))
ggsave(file="Output/Final_Figures/Total-Antioxidant-Capacity.png", TAC.plot, width = 11, height = 6, units = c("in"))
```


Statistical analyses. 
```{r}
```



